---
title: "testJSinHead"
author: "Roger Day"
date: "9/4/2018"
runtime: shiny
output: html_document
---

DISCOVERY:
This doc knits when in misc, 
rmarkdown::run('misc/testJSinHead.Rmd')
JS in a ```{js} section works!
The alert window does appear.
Also works in inst and inst/T15lumpsplit.

---But does not work in project home.
("Error:$ operator is invalid for atomic vectors")

tags$head does not work in Rmarkdown shiny.

textOutput('sessionconnectedID')   NO VALUE

textOutput('safariID')  WORKS

```{js}

  $(document).on('shiny:connected', function(event) {
  alert('shiny:connected  2');
  Shiny.onInputChange('shinyconnected2', 2);
});
```

```{js}
var isSafari = navigator.vendor &&
  navigator.vendor.indexOf('Apple') > -1 &&
  navigator.userAgent &&
  navigator.userAgent.indexOf('CriOS') == -1 &&
  navigator.userAgent.indexOf('FxiOS') == -1;

Shiny.onInputChange("isSafari", isSafari);

if( ! isSafari)
 alert("Please use Safari as your browser for full features.");

```




```{r echo=FALSE}
library(shiny)
require(shinyBS)
require(devtools)
# Define UI for application that draws a histogram

rValues = reactiveValues(temp=1)

output$sessionconnectedID = renderText({
  paste('sessionconnectedID/', input$shinyconnected2, '/')
})

output$safariID = renderText({
  paste('isSafari/', input$isSafari, '/')
})

output$distPlot <- renderPlot({
  # generate bins based on input$bins from ui.R
  x    <- faithful[, 2]
  bins <- seq(min(x), max(x), length.out = input$bins + 1)
  
  # draw the histogram with the specified number of bins
  hist(x, breaks = bins, col = 'darkgray', border = 'white')
})

```

ALL DONE

```{r echo=FALSE}
fluidPage(
  
# Application title
   titlePanel("Old Faithful Geyser Data"),

   # Sidebar with a slider input for number of bins
   sidebarLayout(
      sidebarPanel(
                  textOutput('sessionconnectedID')
      ),

      # Show a plot of the generated distribution
      mainPanel(
         textOutput('safariID')
      )
   )
)
```
