---
title: "Untitled"
author: "Roger Day"
date: "2026-07-21"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#library(shiny)
#options(error=print(traceback()))
#invisible(require(shinyjs) )

```


#  neyman pearson funky demo
# violating conditionality principle


 
```{r echo=FALSE}

h0 = 0 # null
ha = 1 #  alternative hypothesis
vars.init = c(0.4, 3) ####    Small, and large. so power is large, Small.
div(
  numericInput('varsBig', 'varsBig', value = vars.init[2]),
  numericInput('varsSmall', 'varsSmall', value = vars.init[1])
)

```


```{r eval=T}
output[['varsSmall.text']] = renderText(input[['varsSmall']])   # OK.
output[['varsSmall.text.2']] = renderText(input[['varsSmall']])   # OK.

output[['varsBig.text']] = renderText(input[['varsBig']])   # OK.
output[['varsBig.text.2']] = renderText(input[['varsBig']])   # OK.

```

varsSmall.text --   `r textOutput( 'varsSmall.text', inline=TRUE)`
and again...  `r textOutput( 'varsSmall.text', inline=TRUE)`
 The 2nd causes the first to not appear.  
 Overwriting of JS id name.

 and   
varsBig is `r textOutput( 'varsBig.text', inline=TRUE)`  
AND `r textOutput( 'varsBig.text.2', inline=TRUE)` 
Just make more that one output component if needed.
You serialize only as needed.
AND `r textOutput( 'varsBig.text.2', inline=TRUE)`  ???


