# Problemset Example

Example of an RTutor interactive Problemset
Author:  Sebastian Kranz, modified by R.Day
Date:    12.06.2014, now \"`Sys.date()`\"

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.
#< ignore
```{r compare}
### To compare with the original version.
# system(paste0("diff ", "Example_sol.Rmd ",
#    system.file(package='RTutor', 'examples/Example_sol.Rmd')))
```

```{r "create_ps"}
library(restorepoint)
# facilitates error detection
#  I needed to prepend "restorepoint:::"
restorepoint:::set.restore.point.options(display.restore.point=TRUE)

library(RTutor)

#setwd(system.file(package="RTutor", "examples") )
setwd(system.file(package="T15lumpsplit", "examples") )
ps.name = "Example"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = c("ggplot2") # character vector of all packages you load in the problem set
#name.rmd.chunks(sol.file) # set auto chunk names in this file
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, stop.when.finished=FALSE)
show.shiny.ps(ps.name, load.sav=FALSE,  sample.solution=TRUE, is.solved=FALSE, catch.errors=TRUE, launch.browser=TRUE)
stop.without.error()
```
#>

## Exercise 1 -- Summary statistics

a) First part

Some Math:

$$x = \beta^2$$

#< info "Some math in an info block"
$$y = \alpha^2$$
#>


We often want to compute some summary statistic of a vector.
For example:


```{r "1 a)",fig.height=3}
#< task
x = 10:20
# Computing the sum of x
sum(x)
# Generating a data frame
df = data.frame(a=c(1,1,2,2,2,3,3,4), b = runif(8), c="Hi")
#>
#< task_notest
df
#>
```

Now compute the mean of x.
```{r "1 a) 2"}
mean(x)
#< hint
cat("There already exist R functions for many things. To find them use Google, e.g. search for 'R compute mean'.")
#>
```


#< info "useful functions for numeric vectors"
Here are examples for useful R functions
```{r "1 "}
max(c(1,5,2)) # returns maximum
min(c(1,5,2)) # returns minimum

sum(c(1,5,2,NA), na.rm=TRUE) # returns sum of all numbers, ignore NA
cumsum(c(1,5,2)) # returns cummulated sum
diff(c(1,5,2)) # returns the vector of differences
```
#>


#< award "mean means mean"
Well, in some occasions one can just guess the name of an R function. The function to compute the mean of a vector, or matrix is called 'mean'. Usually, it is much quicker to goggle than to guess function names, however.
#>

## Exercise 2 -- Computing with vectors
```{r "2 "}
#< settings
import.var = c("x")
#>
```
a) Let y be a vector that contains the squared elements of x. Show y
```{r "2 a)"}
y = x^2
y
```
