
#### The Bayes Factor: posterior odds (and probability) of the "null hypothesis" <!-- #title -->

The  Bayes Factor is the ratio of these marginal probabilities of the data,  
and also the ratio of posterior odds to prior odds:  

$$
BF = \mathrm{Bayes Factor} = \frac{m_{Split}}{m_{Lump}} = \frac{Pr(Split|data)}{Pr(Lump|data)}
\div \frac{Pr(Split)}{Pr(Lump)}.
$$



```{r , echo=F}
output$BForig = renderText({DrWhoBayesFactor(DLdataOriginal)})
```

```{r , echo=F}
output$BF = renderText({
  analysisName = 'whoMixturePlot'
  source(analysisReactiveSetup_DTC, local=TRUE)
  DrWhoBayesFactor(thisData)
})
```

In our case, the Bayes Factor is `r textOutput('BForig', inline=TRUE)`,  
for the original data, and if you have modified the numbers in the table,  
the Bayes factor is now `r textOutput('BF', inline=TRUE)`.

If *Dr.Split* and *Dr.Lump* have equal prior probability (prior odds = 1),  
then the posterior odds is the Bayes factor times one, 

$$
\rm{odds(}\phi  = Split|data) = \frac{m_{Split}}{{m_{Lump}}} ~~~\times~~~ 1.
$$
which is `r DrWhoBayesFactor(thisData)`  
and  the posterior probability that *Split* is correct is

$$
\rm{Pr(}\phi =Split|data)  = \frac{\rm{odds(}\phi  = Split|data)}
{1~+~\rm{odds(}\phi  = Split|data) } 
= \frac{m_{Split}}{{m_{Split} + {m_{Lump}}}}.
$$
which is `r DrWhoBayesFactor(thisData)/(1+DrWhoBayesFactor(thisData))`  
That's true if the prior odds equals 1  (or 50/50).


```{r , echo=F}
output$bayesAverage = renderText({
  analysisName = 'whoMixturePlot'
  source(analysisReactiveSetup_DTC, local=TRUE)

  paste0('Pr(Lump) *  mLump + Pr(Split) * mSplit',
         '= ',
         "(1/2) * ", mLump(thisData),
         "+
                  (1/2) * ", mSplit(thisData),
         '=',
         signif(digits=3,
                (1/2) * mLump(thisData) +
                  (1/2) * mSplit(thisData))
  )
})

output$postProb = shiny::renderText({round(digits=3,
                                        posteriorProb(priorOdds = 1,
                                                      theData=getDLdata('whoMixturePlot')))})
```
which equals  `r textOutput('bayesAverage', inline=TRUE)`  
for the current data table and Dr.Who's  50% 50% even odds for Lump and Split.

You can adjust the prior probability and observe how  
the posterior probability changes.  
The Bayes factor in favor of  *Dr.Split* is correct is

$$
{\rm{Odds}(\phi = Split|data) = \rm{Odds(}\phi  = Split) \times {m_{Split}}/{m_{Lump}}}.
$$
If Dr. Who puts equal prior probability on *Lump* and *Split*,  
the Bayes factor is the same as the posterior odds.  
Converting from odds to probability,  
with the current data the posterior probability in favor of *Split*  
equals `r  textOutput('postProb', inline=TRUE)`.
