
Dr. Who is not confident in either Dr. Split or Dr. Lump.  Maybe one’s right ($\phi  = Split$ ), maybe the other ( $\phi  = Lump$). Here is a prior that might represent this view:;

\[\pi _{Who}^{}({\bf{p}}) = \pi _{Who}^{}(\phi  = Split)\pi _{Split}^{}({\bf{p}}) + \pi _{Who}^{}(\phi  = Lump)\pi _{Lump}^{}({\bf{p}}).\]
 
NOTE: This marginalizes out (sums over) the nuisance parameter $\phi$ over the joint distribution of ( ${p_{R.D}},\phi$)!  


Briefly Bayes Theorem without a nuisance parameter is (eliminating the usual $\Pr$ & using square brackets, for brevity):

\[\begin{array}{l}
\quad [\theta |X] = \frac{{[X,\theta ]}}{X} = \frac{{[X,\theta ]}}{{\int\limits_\theta  {} [X,\theta ]}} = \frac{{[X|\theta ][\theta ]}}{{\int\limits_\theta  {} [X,\theta ]}}\\
\end{array}.\]

And Bayes Theorem with a nuisance parameter is:

\[\begin{array}{l}
\quad [\theta |X] = \sum\limits_\phi  {} [\theta ,\phi |X] = \frac{{\sum\limits_\phi  {} [X,\theta ,\phi ]}}{{\sum\limits_{\phi ,\theta } {} [X,\theta ,\phi ]}}
\end{array}.\]

Here the nuisance parameter $\phi$ is: whose opinion is right, $Split$ or $Lump$?

More details on Bayes Theorem, including conditioning and marginalizing, are here: 
```{r}
conditionalPanelWithCheckboxPDF(
  labelString="Details on Bayes Theorem", 
  filename='bayes_decision_analysis.pdf',
  cbStringId='bayes_decision_analysis')
```


To apply the Bayes theorem principles:
* We know the data,so we condition on $\bf{n}$. 
* We dont' know -- and don't care about -- $\phi$ ($\phi$ is the "nuisance parameter"), so we "marginalize" over it.
* All that's left is $\bf{p}$, what we care about.

Then,  Dr. Who’s posterior distribution for ${\bf{p}}$ is:
 \[\pi _{Who}^*({p_{R.D}}) = \pi _{Who}^*(\phi  = Split)\pi _{Split}^*({p_{R.D}}) + \pi _{Who}^*(\phi  = Lump)\pi _{Lump}^*({p_{R.D}}),\]

where
 \[\pi _{Who}^*(\phi  = Split) = \frac{{\pi (Split){m_{Split}}}}{{\pi (Split){m_{Split}} + \pi (Lump){m_{Lump}}}},\]
and the $m$ values are the respective marginal distributions of the data  $({n_{RD}},{n_{RL}},{n_{ND}},{n_{NL}})$ if you knew $\phi$ but not the parameters  --  conditional $\phi$ and marginal over ${p_{R.D}}$.  

```{r}
TextQuestion("If Dr. Who gives Dr. Lump and Dr. Split equal weight BEFORE seeing the data, is this still true AFTER seeing the data?
")
```

#### Marginal probability for each hypothesis Split and Lump <!-- #title -->

```{r}
output$mSplit = renderText({
  signif(digits=3, mSplit(rValues$DLdata)); 
  })
```

If *Split* is correct, then the (marginal) probability of getting the data we observed is 
**`r   textOutput('mSplit', inline=TRUE)`**.

For the technical details, take a look here:
```{r}
conditionalPanelWithCheckbox(
  labelString = 'Equations for Dr. Split\'s marginal probability', 
  filename = 'DrSplitEquation.Rmd',
  html='',
  initialValue=FALSE
)
```

```{r}
output$mLump = renderText({
  signif(digits=3, mLump(rValues$DLdata)); 
  })
```

If *Lump* is correct, then the (marginal) probability of getting the data we observed is:
**`r   textOutput('mLump', inline=TRUE)`**.

For the technical details, take a look here:
```{r}
conditionalPanelWithCheckbox(
  labelString = 'Equations for Dr. Lump\'s marginal probability', 
  filename = 'DrLumpEquation.Rmd',
  html='',
  initialValue=FALSE
)
```

Here is the code to calculating the marginals and the Bayes factor if you are interested:

```{r}
DrWhoBayesFactor.code = readLines(con='DrWhoBayesFactor.R')
DrWhoBayesFactor.code = 
 HTML(gsub('  ', '&nbsp;&nbsp;',
                        paste(DrWhoBayesFactor.code,
                    collapse='<br>\n')))
conditionalPanelWithCheckbox(
  labelString = 'Code for the Lump and Split marginal probabilities',
  html=DrWhoBayesFactor.code,
  initialValue=FALSE
)
```

The posterior distribution for Dr. Who, which marginalizes over $\phi$ as we have seen, is a mixture, as graphed here.

Viewing Dr. Who's posterior:

```{r}
jumpList = c(jumpList, whoMixturePlot='Posterior distribution for Dr. Who')
source('analysisInitialSetup.R', local=TRUE)

output$whoMixturePlot = 
  renderPlot(expr =  {
    analysisName = 'whoMixturePlot'
    source('analysisReactiveSetup.R', local=TRUE)
    pSeq = seq(0,1, length=100)
    par(mfrow=c(1,2))
    prior.s = dbeta(pSeq, 1, 1)
    prior.l = dbeta(pSeq, 2, 2)
    prior.w = (1/2) * prior.s + (1/2) * prior.l
    plot(pSeq, prior.s, col='blue', type='l',
         xlab='probability', ylab='density',
         ylim=c(0,15),
         main='Priors') 
    lines(pSeq, prior.l,  col='red')
    lines(pSeq, prior.w,  col='darkgreen', lwd=2)
    everyTenth = (1:length(pSeq)) %% 10 == 0
    points(pSeq[everyTenth], 
           prior.s[everyTenth], pch='S', col='blue')
    points(pSeq[everyTenth], 
           prior.l[everyTenth], pch='L', col='red')
    points(pSeq[everyTenth], 
           prior.w[everyTenth], pch='W', col='darkgreen')
    Sexpression = as.expression(
      bquote("S:  " * italic(
        p[R.D]) %~% italic(Be*ta)(2, 2) )  )
    text(0, 11, labels = Sexpression,
         col='blue', cex=1.5, pos=4
    )
    Lexpression = as.expression(
      bquote( "L:  " * italic(
        p[R.D]==p[R]) %~% Be*ta(2, 2) ) ) 
    text(0, 9, labels = Lexpression,
         col='red', cex=1.5, pos=4
    )
    Wexpression = as.expression(
      bquote( "W:  " * italic("Dr. Who mixture" ) ) )
    text(0, 7, labels = Wexpression,
         col='darkgreen', cex=1.5, pos=4
    )
#currentTableOfEstimates = tableOfEstimates()
    nRD=rValues$DLdata['R','D']
    nND=rValues$DLdata['N','D']
    posterior.s = dbeta(pSeq, 
                    1+nRD, 
                    1+nND)
    posterior.mean.s = (1+nRD)/(2+nRD+nND)
    nR=sum(rValues$DLdata['R', ])
    nN=sum(rValues$DLdata['N', ])
    posterior.mean.l = (2+nR)/(4+nN)
    posterior.l = dbeta(pSeq, 
                        2+nR, 2+nN)
    
    #cat('mSplit is ', mSplit(rValues$DLdata), '\n')
    #cat('mLump is ', mLump(rValues$DLdata), '\n')
    #cat('DrWhoBayesFactor is ', DrWhoBayesFactor(rValues$DLdata), '\n')
    pProbSplit = posteriorProb(theData=rValues$DLdata)
    pProbLump = (1-pProbSplit)
    posterior.w = pProbSplit * posterior.s + pProbLump * posterior.l
    posterior.mean.w = 
      pProbSplit * posterior.mean.s +
      pProbLump * posterior.mean.l
    plot(pSeq, posterior.s, col='blue', type='l',
         xlab='probability', ylab='density',
         ylim=c(0,15),
         main='Posterior densities and means') 
    lines(pSeq, posterior.l,  col='red')
    lines(pSeq, posterior.w,  col='darkgreen', lwd=2)

    points(pSeq[everyTenth], posterior.s[everyTenth], 
           pch='S', col='blue')
    points(pSeq[everyTenth], posterior.l[everyTenth], 
           pch='L', col='red')
    points(pSeq[everyTenth], posterior.w[everyTenth], 
           pch='W', col='darkgreen')

    offset = 0.2
    Sexpression = as.expression(
      bquote("S:  " * italic(
        p[R.D]) %~% italic(Be*ta)(2+.(nRD), 2+.(nND))) ) 
    text(offset+0,11, labels = Sexpression,
         col='blue', cex=1.5, pos=4, xpd=NA
    )
    Lexpression = as.expression(
      bquote( "L:  " * italic(
        p[R.D]==p[R]) %~% Be*ta(2+.(nR), 2+.(nN))) ) 
    text(offset+0, 9, labels = Lexpression,
         col='red', cex=1.5, pos=4, xpd=NA
    )
    Wexpression = as.expression(
      bquote( "W:  " * italic("Dr. Who mixture" ) ) )
    text(offset+0, 7, labels = Wexpression,
         col='darkgreen', cex=1.5, pos=4, xpd=NA
    )
    text(offset+0.2, 5, labels = 
           paste0(round(digits=2, pProbSplit)),
         col='darkgreen', cex=1.5, pos=4, xpd=NA
    )
    text(offset+0.35, 5, labels = "S",
         col='blue', cex=1.5, pos=4, xpd=NA)
    text(offset+0.45, 5, labels = "+",
         col='darkgreen', cex=1.5, pos=4, xpd=NA)
    text(offset+0.5, 5, labels = 
           paste0(round(digits=2, pProbLump)),
         col='darkgreen', cex=1.5, pos=4, xpd=NA
    )
    text(offset+0.65, 5, labels = "L",
         col='red', cex=1.5, pos=4)
    onTop = par('usr')[4]  ## ['y2']
    #cat('onTop: ', onTop, '\n')
    ######  Mark the positions of the means.
    text(0, onTop*1.05, col='black', xpd=NA, 
         labels=expression(Means %=>% phantom('.')), cex=1, pos = 2, font=4)
    text(posterior.mean.l, onTop*1.00, col='red', xpd=NA, labels="L", cex=1, pos = 3, font=4)
    text(posterior.mean.s, onTop*1.00, col='blue', xpd=NA, labels="S", cex=1, pos = 3, font=4)
    text(posterior.mean.w, onTop*1.00, col='darkgreen', xpd=NA, labels="W", cex=1, pos = 3, font=4)
  })
plotOutput('whoMixturePlot')
dataTableComponent(showhide='hide', analysisName='whoMixturePlot')
TextQuestion("Experiment with changing #RL in the range 2 to 9. What happens?  And... why? ")
QandAha('QA-whoMixturePlot')
```

#### The Bayes Factor: posterior odds (and probability) of the "null hypothesis" <!-- #title -->

The  Bayes Factor is the ratio of these marginal probabilities of the data,
and also the ratio of posterior odds to prior odds:
\[BF = \frac{m_{Split}}{m_{Lump}} = \frac{Pr(Split|data)}{Pr(Lump|data)}
\div \frac{Pr(Split)}{Pr(Lump)}.\]

```{r}
output$BForig = renderText({DrWhoBayesFactor(DLdataOriginal)})
```

```{r}
output$BF = renderText({DrWhoBayesFactor(rValues$DLdata)})
```

In our case, the Bayes Factor is `r textOutput('BForig', inline=TRUE)`, for the original data, and if you have modified the numbers in the table, the Bayes factor is now `r textOutput('BF', inline=TRUE)`.

If *Dr.Split* and *Dr.Lump* have equal prior probability (prior odds = 1),
then the posterior odds is the Bayes factor,
and  the posterior probability that *Dr.Lump* is correct is

\[{\rm{Pr(}}\phi  = Lump|data) = {m_{Lump}}/({m_{Split}} + {m_{Lump}}).\]

More generally, 

\[
\Pr (R|D,Who) = \Pr (Lump|data) \times \Pr (R|D,Lump) + \Pr (Split|data) \times \Pr (R|D,Split),  \]

```{r}
output$bayesAverage = renderText({
  paste0('Pr(Lump) *  mLump + Pr(Split) * mSplit',
         '= ',
         signif(digits=3,
                (1/2) * mLump(rValues$DLdata) +
                  (1/2) * mSplit(rValues$DLdata))
  )
})
```
which equals  `r textOutput('bayesAverage', inline=TRUE)` for the current data table.

You can adjust the prior probability and observe how the posterior probability changes.
The Bayes factor in favor of  *Dr.Split* is correct is

\[{\rm{Odds}(\phi = Split|data) = \rm{Odds(}\phi  = Split) \times {m_{Split}}/{m_{Lump}}}.\]
If Dr. Who puts equal prior probability on *Lump* and *Split*, the Bayes factor is the same as the posterior odds. Converting from odds to probability, with the current datathe posterior probability in favor of *Split* equals `r output$postProb = shiny::renderText({round(digits=3, posteriorProb(theData=rValues$DLdata))}); textOutput('postProb', inline=TRUE)`.
