
Here is a prior that might represent this view:
  
  
$$
\def\Who{\mathbf{Who}}
\begin{aligned}
\pi _{\Who}^{}({\bf{p}}) = & ~~~~~\pi _{\Who}^{}(\phi  = Split)~~~\pi _{Split}^{}({\bf{p}}) \\
& + \pi _{\Who}^{}(\phi  = Lump)~~\pi _{Lump}^{}({\bf{p}}).
\end{aligned}
$$
  
  NOTE: This marginalizes out (sums over)  
the *nuisance parameter* $\phi$  (who is right?)  
over the joint distribution of ( ${p_{R.D}},\phi$)!  
  
  
  Briefly Bayes Theorem without a nuisance parameter is  
(eliminating the usual $\Pr$ & using square brackets, for brevity):
  
$$
\begin{array}{l}
    \quad [\theta |X] = \frac{{[X,\theta ]}}{X} = \frac{{[X,\theta ]}}{{\int\limits_\theta  {} [X,\theta ]}} = \frac{{[X|\theta ][\theta ]}}{{\int\limits_\theta  {} [X,\theta ]}}\\
    \end{array}.
$$

And Bayes Theorem with a nuisance parameter is:
  
$$
\begin{array}{l}
    \quad [\theta |X] = \sum\limits_\phi  {} [\theta ,\phi |X] = \frac{{\sum\limits_\phi  {} [X,\theta ,\phi ]}}{{\sum\limits_{\phi ,\theta } {} [X,\theta ,\phi ]}}
    \end{array}.
$$

Here the nuisance parameter $\phi$ is:  
  whose opinion is right, $Split$ or $Lump$?
  
  More details on Bayes Theorem,  
including conditioning and marginalizing, are here: 
```{r}
conditionalPanelWithCheckboxPDF(
  labelString="Details on Bayes Theorem", 
  filename='bayes_decision_analysis.pdf',
  cbStringId='bayes_decision_analysis')
```


To apply the Bayes theorem principles:  
  * We know the data, so we condition on $\bf{n}$.   
* We dont' know (and don't care about) $\phi$  
  ($\phi$ is the "nuisance parameter"), so we "marginalize" over it.  
* All that's left is $\bf{p}$, what we care about.

Then,  Dr. Who’s posterior distribution for ${\bf{p}}$ is:
$$\pi _{\Who}^*({p_{R.D}}) = \pi _{\Who}^*(\phi  = Split)\pi _{Split}^*({p_{R.D}})  
 + \pi _{\Who}^*(\phi  = Lump)\pi _{Lump}^*({p_{R.D}}),
$$

where
$$\pi _{\Who}^*(\phi  = Split) = \frac{{\pi (Split){m_{Split}}}}{{\pi (Split){m_{Split}} + \pi (Lump){m_{Lump}}}},
$$
and the $m$ values are the respective marginal distributions of the data   $({n_{RD}},{n_{RL}},{n_{ND}},{n_{NL}})$  

if you knew $\phi$ but not the parameters   
-- then you condition $\phi$ and marginalize over ${p_{R.D}}$  
to get what you care about,  
${p_{R|D}}$ 

```{r}
TextQuestion("If Dr. Who gives Dr. Lump and Dr. Split equal weight BEFORE seeing the data, is this still true AFTER seeing the data?
")
```

#### Marginal probability for each hypothesis Split and Lump <!-- #title -->

```{r, echo=FALSE}
output$mSplit = renderText({
  analysisName = 'whoMixturePlot'
  source(analysisReactiveSetup_DTC, local=TRUE)
  signif(digits=3, mSplit(thisData)); 
  })
```

If *Split* is correct, then the (marginal) probability of getting the data we observed is  

**`r   textOutput('mSplit', inline=TRUE)`**.

For the technical details, take a look here:

```{r sidebars, echo =F}
## nice try, doesn't work, the input name not found..
sidebar = function(RmdFileName){
  h4content = paste0("Sidebar:  `r actionButton(class='sidebar', inputId=",
            RmdFileName,
           "', label =  ", RmdFileName, " )`  <!-- #title -->"
  )
  print(h4content)
  h4(h4content) 
  observe( {
    if(!is.null(input[[RmdFileName]]) ) {
      if(input[[RmdFileName]] > 0) {
        showModal(modalDialog(size='xl', title=RmdFileName, easyClose = T,
                              footer=modalButton('OK'),
                              inclRmd(RmdFileName)))
      }
    }
  })
}
```

#### : Sidebar:  `r actionButton(class='sidebar', inputId='DrSplitEquation.Rmd', label =  'DrSplitEquation.Rmd' )`  <!-- #title -->
```{r}
#### this works!
observe( {
  if(!is.null(input$DrSplitEquation.Rmd) ) {
    if(input$DrSplitEquation.Rmd > 0) {
        showModal(modalDialog(size='xl', title="DrSplitEquation.Rmd", easyClose = T,
                               footer=modalButton('OK'),
                              inclRmd('DrSplitEquation.Rmd')))
    }
  }
}
)
```



```{r}
output$mLump = renderText({
  analysisName = 'whoMixturePlot'
  source(analysisReactiveSetup_DTC, local=TRUE)
  signif(digits=3, mLump(thisData)); 
})
```

If *Lump* is correct,  
then the (marginal) probability of getting the data we observed is:
  **`r   textOutput('mLump', inline=TRUE)`**.

For the technical details, take a look here:
#### : Sidebar:  `r actionButton(class='sidebar', inputId='DrLumpEquation.Rmd', label =  'DrLumpEquation.Rmd' )`  <!-- #title -->
```{r}
#sidebar('DrLumpEquation.Rmd') ### too bad doesn't work.

#### this works!
observe( {
  if(!is.null(input$DrLumpEquation.Rmd) ) {
    if(input$DrLumpEquation.Rmd > 0) {
        showModal(modalDialog(size='xl', title="DrLumpEquation.Rmd", easyClose = T,
                               footer=modalButton('OK'),
                              inclRmd('DrLumpEquation.Rmd')))
    }
  }
}
)
```

Here is the code to calculating the marginals and the Bayes factor  
if you are interested:
  
```{r}
DrWhoBayesFactor.code = readLines(con='DrWhoBayesFactor.R')
DrWhoBayesFactor.code = 
  HTML(gsub('  ', '&nbsp;&nbsp;',
            paste(DrWhoBayesFactor.code,
                  collapse='<br>\n')))
conditionalPanelWithCheckbox(
  labelString = 'Code for the Lump and Split marginal probabilities',
  html=DrWhoBayesFactor.code,
  initialValue=FALSE
)
```
