
 To deploy:
 The most recent installation of the package must be with
 .installFromGithub()
 
Then, select the tab with the Rmd file,
and use the blue broken circle button to Publish to shinyapps.io.
(See my "NOTES deploying with bioconductor package".  For me, qvalue.)

To view logs of the shinyapps run, go to https://www.shinyapps.io/admin ,
select the running app, and click the log tab.

Child chunks may not have chunk names or content, not even a comment.
Installing from github you get a warning.
But then deploying bombs with a mysterious message, can't find the temp file path.

```{r}
# options(repos = c(BiocInstaller::biocinstallRepos()  
#                   , "https://cran.rstudio.com/") )
## options(repos = c(BiocInstaller::biocinstallRepos()  ) )
## "https://cran.rstudio.com/" is redundant 

rsconnect::appDependencies(appDir = 'inst/T15lumpsplit/')
.install...
rsconnect::showLogs(appPath = 'inst/T15lumpsplit')

```

Works fine at shinyapps.io, but an error in browserAdvice.
See log file at https://www.shinyapps.io/admin/#/application/398397/logs 
Warning: Error in if: argument is of length zero
2018-10-06T03:40:49.594543+00:00 shinyapps[398397]:   106: renderUI [<text>#9]
2018-10-06T03:40:49.594544+00:00 shinyapps[398397]:   105: func
2018-10-06T03:40:49.594546+00:00 shinyapps[398397]:    92: origRenderFunc
2018-10-06T03:40:49.594546+00:00 shinyapps[398397]:    91: output$browserAdvice


```{r include navTOCid}
#source('navTOCid.R', local=T)
```

```{r}
includeScript('www/browserDetect.js')
```

