---
title: "Untitled"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
```


#  neyman pearson funky demo
# violating conditionality principle



```{r echo=FALSE}

h0 = 0  # null
ha = 1 #  alternative hypothesis
vars = c(0.4, 3) ####    Small, and large. so power is large, Small.
alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
  lowerC = 0
  Cseq <- seq(lowerC,3,length=1000)

likelihood = function(h, v, x) {
   1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
# likelihood(h = 0, v = 1, x = qnorm(1-0.025)) # yes 0.025.
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
```

```{r eval=FALSE}

### skip this plot
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))
abline(v=0:1)
lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[2])))
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))

```

The setting:  one observation from a normal distribution, mean unknown.
If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
```{r echo=FALSE}
  cutoffs  = qnorm(1 - alpha, sd = sqrt(vars))
lrEqualAlpha =  lr(cutoffs, vars)
  pow.M.alt.1 = pow(cutoffs, M.alt=1, V=vars)
  results.EqualAlpha = data.frame(
    alphas = rep(alpha, 2),
    cutoffs, 
    LR= lrEqualAlpha, 
    power=pow.M.alt.1,
    marginalPower= rep(mean(pow.M.alt.1), 2))

```
But we have **two** instruments to measure *X*, with different variances  `r vars[1]` and `r vars[2]`.  They could be different in precisions. Or they could be because there are two different possible sample sizes.  *The instrument used is picked at random!*  ---- **But we know which one was picked!**

Suppose we use the same *alpha*, the same conditional type I error rate, regardless of which instrument. If you know which instrument was picked, you would of course use the relevant significance cutoff. Here for *alpha* = `r alpha`, used regardless of the instrument chosen, these two conditional cutoffs are `r signif(digits=4, cutoffs[1])` and `r signif(digits=4, cutoffs[2])`.  Then regardless of the chances for the two instrument, the overall marginal *alpha* is the same.

The likelihood ratios at the two critical values are `r lrEqualAlpha[1]` and `r lrEqualAlpha[2]`.  Not the same.  Should we maximize power by  forcing the two likelihood ratios to be the same?

If we raise the cutoff for the Bigger variance `r vars[1]`, and lower it for the Smaller variance `r vars[2]`, watch what happens. The likelihood ratios get closer and closer.

```{r}
Aseq<- seq(0.001, 0.05, length=1000)
lrBigValues = numeric(0)
lrSmallValues = numeric(0)
cutoffsBigValues = numeric(0)
cutoffsSmallValues = numeric(0)
lrDifference = function(alphaBig){
  alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
  cutoffsBig = qnorm(1 - alphaBig, sd = sqrt(max(vars)))
  cutoffsSmall = qnorm(1 - alphaSmall, sd = sqrt(min(vars)))
  cutoffsBigValues =  append(cutoffsBigValues, cutoffsBig)
  cutoffsSmallValues =  append(cutoffsSmallValues, cutoffsSmall)
  #print(c( lr(cutoffsBig, v = max(vars)) , lr(cutoffsSmall, v = min(vars)) ) )
  lrBig = lr(cutoffsBig, v = max(vars)) 
  lrSmall = lr(cutoffsSmall, v = min(vars))
  lrBigValues =  append(lrBigValues, lrBig)
  lrSmallValues =  append(lrSmallValues, lrSmall)
  lrBig - lrSmall
}
lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
optimal = optimize(lrDifferenceAbs, 
                           maximum = FALSE, interval = c(0, max(Aseq))) 
optimalAlphaBig = optimal$ minimum
whichOptimal = ( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
optimalAlphaSmall = 2*alpha - optimalAlphaBig
cutoffsBig =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars)))
cutoffsSmall = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars)))
# testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
#                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars[1]), 
                               pow(cutoffsBig, M.alt=1, V=vars[2]) )

averagePowerEqualLR = mean(conditionalPowerEqualLR)
  
results.EqualLR <<- data.frame(
  alphas = c(optimalAlphaSmall, optimalAlphaBig), 
  cutoffs = c(cutoffsSmall, cutoffsBig), 
  LR = rep(optimal$objective, 2),
  power = conditionalPowerEqualLR, 
  marginalPower = rep(averagePowerEqualLR, 2))

output$lrDifferenceAbs = renderPlot({
  plot(Aseq,
     lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
  abline(v=alpha/2, lty=2)
  abline(v=optimalAlphaBig, lty=3)
})
plotOutput('lrDifferenceAbs')

```

The plot shows the absolute difference in the two likelihood ratios, as we change the proportion of **alpha** for the big variance. We see that the likelihood ratios are equal when the alpha for the big variance instrument is **`r round(optimalAlphaSmall, digits=3)`**. Then the other **alpha**, for the less precise instrument, is `r round(optimalAlphaBig, digits=3)`. They average to **alpha = `r alpha`**.

```{r echo=FALSE}
### conditional power

output$theplot = renderPlot({
  plot(Cseq, ylim=c(0,1), xlim=c(0, cutoffsBig * 1.01),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
  lines(x=c(lowerC, cutoffs[1]), y= c(alpha,alpha) , col='black')
  lines(x=c(lowerC, cutoffs[2]), y= c(alpha,alpha) , col='black')
  text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  points(cutoffs, c(alpha, alpha), col=col.v, pch='C')
  #points(cutoffs, pow(cutoffs, M.alt=0), col=col.v, pch='C')
  points(cutoffs, pow.M.alt.1, col=col.v, pch='P')
  
  
  for(Vn in 1:2) {
    lines(Cseq, pow(Cseq, M.alt=1, V=vars[Vn]), lty=2, lwd=3, col=col.v[Vn])
    lines(x=c(lowerC, cutoffs[Vn]), y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
    lines(x=rep(cutoffs[Vn],2), y = c(alpha, pow.M.alt.1 [Vn]), col=col.v[Vn])
    text(x=lowerC, y=pow.M.alt.1 [Vn], col=col.v[Vn],
         labels = paste('V=', vars[Vn], '  power=', 
                        signif(digits = 2, pow.M.alt.1 [Vn])), adj = c(0,0))
    
  }
  ### average
  averagePowerEqualAlpha = mean(pow.M.alt.1)
#  lines(x=c(lowerC, mean(cutoffs)), y= rep(averagePowerEqualAlpha, 2), col=col.v[3])
  lines(rep(lowerC, 2), pow.M.alt.1, col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal alpha:', 
                      signif(digits = 6, averagePowerEqualAlpha)))

  
  points( cex=2, c(cutoffsSmall, cutoffsBig), 
          conditionalPowerEqualLR, col=col.v[4], pch='L')
  text(x=lowerC, y=averagePowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v[4], xpd=NA,
       labels = paste('marginal power equal LR:', 
                      signif(digits = 6, averagePowerEqualLR)))
  
  
  legend(x='topright', 
         legend=c(paste('Variance=', vars[1]), 
                  paste('Variance=', vars[2]), 
                  'equal alpha',
                  'equal LR'
         ), 
         text.col =  col.v
  )
})


plotOutput('theplot')
```
```{r eval=FALSE}
output$likrat = renderPlot({
  plot(c(cutoffsSmallValues, cutoffsBigValues), 
       )
})
```



If the variance is `r vars[1]`, the cutoff for alpha=0.05 is `r signif(digits=3, cutoffs[1] )`.  
But if the variance is `r vars[2]`, the cutoff for alpha=0.05 is `r signif(digits=3, cutoffs[2] ) `.  
The conditional statistical **powers** for rejecting the "null" (bland) hypothesis $M=0$  are `r pow.M.alt.1[1]` and `r pow.M.alt.1[2]`,


and the average power, averaging the conditional powers
over  which variance is chosen (50-50), is 

## Likelihood ratios
The likelihood ratios are, respectively, 

`r lr(cutoffs[1], vars[1])` and `r lr(cutoffs[2], vars[2])`

Not equal!!

The equalizer for the likelihood ratio is when  
AlphaBig = optimalAlphaBig = `r optimalAlphaBig` and  
AlphaSmall = optimalAlphaSmall = `r optimalAlphaSmall`.  
The critical value for the small variance is now `r cutoffsSmall`,
and for the large variance it's `r cutoffsBig`.
```{r}
    results.EqualLR = data.frame(
      alphas = c(optimalAlphaBig, optimalAlphaSmall),
      cutoffs=c(cutoffsBig, cutoffsSmall), 
                                 lrEqualAlpha, pow.M.alt.1)

  

```
The increase in power is minuscule.  
But you are actually combining two possible experiments and ignoring the fact that you know which one was done.

Compare the cutoffs.  
With equal *alpha*:  
Are you content to report a P value based on the marginal **alpha** when you KNOW which instrument was used? So the **conditionality p(rinciple** is used FIRST.

```{r}
  output$results.EqualAlpha = renderTable(results.EqualAlpha )
  tableOutput('results.EqualAlpha')
```

```{r}  
  output$results.EqualLR = renderTable(results.EqualLR) 
  tableOutput('results.EqualLR')
```




