---
title: "neyman pearson funky demo"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
resource_files:
- www/KeyHandler.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(DT)  # not data.table
library(reactable)
library(shinyDebuggingPanel, quietly = TRUE)
options(error=print(traceback()))
options(shiny.launch.browser=TRUE)


tO = function(V) textOutput(inline=TRUE, V)
```


#  The beautiful Neyman Pearson lemma makes you do nutty things



```{r echo=FALSE}
#shiny::includeScript(path =  paste0(getwd(), '/inst/T15lumpsplit/www/KeyHandler.js') )  FAILS--- private folder

includeScript('/Users/rogerday/Google Drive/Documents/T15lumpsplit/inst/T15lumpsplit/www/KeyHandler.js') 
observeEvent(input$ctrlDpressed, {}) # just to flush the ctrl-D press.

shinyDebuggingPanel::makeDebuggingPanelOutput(
   session, toolsInitialState = TRUE,
   condition='ctrlDpressed === true')

fluidRow(
shinyDebuggingPanel::withDebuggingPanel()
)



```

```{r echo=FALSE}
h0 = 0 # null
ha = 1 #  alternative hypothesis
vars.init = c(small=0.4, big=3) ####    Small, and large. so power is large, Small.
vars.step = 0.1

vars = reactive({c(input$varsSmall, input$varsBig)})
observeEvent(input$varsBig,
             updateNumericInput(inputId='varsSmall', max = input$varsBig))
observeEvent(input$varsSmall,
             updateNumericInput(inputId='varsBig', min = input$varsSmall))

alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
lowerC = 0

likelihood = function(h, v, x) {
  1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
# likelihood(h = 0, v = 1, x = qnorm(1-0.025)) # yes 0.025.
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
```

```{r eval=FALSE}

### skip this plot
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall())))
# abline(v=0:1)
# lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsBig() )))
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall() )))

```

The setting:  one observation from a normal distribution, mean unknown.
If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
```{r echo=FALSE}
output[['varsSmall.text']] = renderText(input[['varsSmall']])   # OK.
output[['varsSmall.text.2']] = renderText(input[['varsSmall']])   # OK.

output[['varsBig.text']] = renderText(input[['varsBig']])   # OK.
output[['varsBig.text.2']] = renderText(input[['varsBig']])   # OK.

cutoffs.by.alpha = function(alphaBoth, variancesBoth)
    qnorm(1 - alphaBoth, sd = sqrt(variancesBoth))


cutoffs.equalAlpha = reactive({
  cutoffs.by.alpha(c(alpha, alpha), vars())
})
cutoffs.equalAlpha_1 = reactive({
  cutoffs.by.alpha(alpha, vars()[1])
})
cutoffs.equalAlpha_2 = reactive({
  cutoffs.by.alpha(alpha, vars()[2])
})


output$cutoffsAlpha.text = renderText(paste0(signif(digits=3,
                                                    cutoffs.equalAlpha()[1]), 
                                             ' and ',
                                             signif(digits=3, cutoffs.equalAlpha()[2] ) ) )
output$cutoffsAlpha1.text = renderText(paste0(signif(digits=3, cutoffs.equalAlpha_1)))
output$cutoffsAlpha2.text = renderText(paste0(signif(digits=3, cutoffs.equalAlpha_2)))

results.EqualAlpha.Table = reactive({
    #cutoffs  = qnorm(1 - alpha, sd = sqrt(vars))
  lrEqualAlpha <<-  lr(cutoffs.equalAlpha(), vars())
  pow.M.alt.1 <<- pow(cutoffs.equalAlpha(), M.alt=1, V=vars())
  results.EqualAlpha <<- #data.frame(
    #data = 
#      signif(digits=2, 
    data.frame(
        alphas = rep(alpha, 2),
        cutoffs=cutoffs.equalAlpha(), 
        LR= lrEqualAlpha, 
        power=pow.M.alt.1,
        marginalPower= rep(mean(pow.M.alt.1), 2)
      )
    for(i in ncol(results.EqualAlpha))
      results.EqualAlpha[[i]] = signif( digits=3, results.EqualAlpha[[i]])
  # newNames = colDef(names(results.EqualAlpha))
  # print(newNames)
  # newNames[[5]] = colDef('marginal<br>power', html=TRUE)
  # print(newNames)
  # names(newNames = names(results.EqualAlpha))
  # print(newNames)
  (results.EqualAlpha )
            #, columns = newNames )
  })

output$results.EqualAlpha = renderTable({
  df = results.EqualAlpha.Table()
  names(df)[names(df) == 'marginalPower'] = 'marginal<br>power'
  names(df)[names(df) == 'power'] = 'conditional<br>power'
  df
}, sanitize.colnames.function = function(x) x)


```
But we have **two** instruments to measure *X*, with different variances
**`r textOutput(inline=T, 'varsSmall.text')` and `r textOutput(inline=T, 'varsBig.text')`.**  
They could be different in precisions. Or they could be because there are two different possible sample sizes. 

*The instrument used is picked at random!*  

---- **And...  we know which one was picked!**

Suppose we use the same *alpha*, the same conditional type I error rate, regardless of which instrument. We'll call that  
**"equal alpha"**.  
If you know which instrument was picked, you will of course use the relevant significance cutoff. Here for *alpha* = `r alpha`, used regardless of the instrument chosen, these two conditional cutoffs are `r textOutput(inline=T, 'cutoffs.text')`.  
Then regardless of the chances for selection either instrument, the overall **marginal** *alpha* is the same.
```{r}
# lrEqualAlpha <<-  lr(cutoffsAlpha(), vars())

# pow.M.alt.1 <<- reactive(pow(cutoffsAlpha(), M.alt=1, V=vars()) )

output$lrEqualAlpha.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha.Table()$LR[1]),
         ' and ',
         signif(digits=3, results.EqualAlpha.Table()$LR[2])
         )
})
output$lrEqualAlpha_1.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha.Table()$LR[1]) )
})
output$lrEqualAlpha_2.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha.Table()$LR[2]) )
})

```

But why not maximize the statistical power?  
Our null  hypothesis is  **mean = 0**.  
Pick the alternative hypothesis which we want power against to be **mean = 1**.  

```{r}
div(
  numericInput('varsBig', 'variance Big', value = vars.init[2], min = vars.init[1], step=vars.step),
  numericInput('varsSmall', 'variance Small', value = vars.init[1], min = 0,
               max = vars.init[2], step=vars.step)
)
```

Our current rejection region is  
```{r}
output$rejectionRegion.text = renderUI(
  HTML(paste0(
    span(style='color:darkgreen', 
         'variance = ', vars()[2],
         " AND   cutoff = ", 
         signif(digits=3,
                cutoffs.equalAlpha_2())
    ), 
         ' OR ',
    span(style='color:red', 
         'variance = ', vars()[1],
         " AND   cutoff = ", 
         signif(digits=3, cutoffs.equalAlpha_1()
         ) )
  )
))
    
div(style='text-align:center',
    uiOutput('rejectionRegion.text')
)
```


```{r plotRejectionRegion}
output$plotRejectionRegion = renderPlot( height=200, #width='100%', 
                                        {
  print(vars.init)
  plot.default( xlab= 'observed value of X',
                ylab = 'variance',
                xlim=c(0, results.EqualAlpha.Table()$cutoffs[2] * 2),
                ylim = vars.init+ c( -0.1, 0.1),
    axes=F,
    c( 0, results.EqualAlpha.Table()$cutoffs[2] * 2), 
       vars.init+ c( -0.1, 0.1), col=NA) 
#  abline(h=vars.init)  
  for (v in 1:2) {
    lines(c(0, results.EqualAlpha.Table()$cutoffs[v]), rep(vars()[v], 2),
          col = col.v[v])
    lines(c(results.EqualAlpha.Table()$cutoffs[v],
            results.EqualAlpha.Table()$cutoffs[2] * 2),  
          rep(vars()[v], each=2) , col = col.v[v],
          lwd=4)
  }
  points( results.EqualAlpha.Table()$cutoffs, 
          vars.init, col=col.v[1:2], cex=2, pch=1) 
  text(results.EqualAlpha.Table()$cutoffs[2], vars()[2],
       'rejection region, var = 3.0' ,  adj=c(-1,-1), xpd=NA)
  text(results.EqualAlpha.Table()$cutoffs[1], vars()[1],
       'rejection region, var = 0.4',   adj=c(-1,-1))
  text(0.05, vars()[1] - 0.3, '0', xpd=NA)
} )
plotOutput('plotRejectionRegion', height = 200)
```


The likelihood ratios at the two critical values are `r textOutput(inline=TRUE, 'lrEqualAlpha.text')`.
Not the same.   
The "Neyman-Pearson lemma" says, maximize power by setting the rejection region boundary to have the same likelihood ratio everywhere on the boundary.
Should we maximize power by  forcing the two likelihood ratios to be the same?

If we raise the cutoff for the Bigger variance `r textOutput(inline=T, 'varsBig.text.2')`, and lower it for the Smaller variance `r textOutput(inline=T, 'varsSmall.text.2')`, watch what happens. The likelihood ratios get closer and closer.

```{r}
  lrDifference = function(alphaBig){
      lrBigValues = numeric(0)
      lrSmallValues = numeric(0)
      cutoffsBigValues = numeric(0)
      cutoffsSmallValues = numeric(0)
      
    alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
    cutoffsBig = qnorm(1 - alphaBig, sd = sqrt(max(vars())))
    cutoffsSmall = qnorm(1 - alphaSmall, sd = sqrt(min(vars())))
    cutoffsBigValues =  append(cutoffsBigValues, cutoffsBig)
    cutoffsSmallValues =  append(cutoffsSmallValues, cutoffsSmall)
    #print(c( lr(cutoffsBig, v = max(vars)) , lr(cutoffsSmall, v = min(vars)) ) )
    lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    lrBigValues =  append(lrBigValues, lrBig)
    lrSmallValues =  append(lrSmallValues, lrSmall)
    lrBig - lrSmall
  }
  lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
```



```{r}
results.EqualLR.Table = reactive({
  Aseq<- seq(0.001, 0.05, length=1000)
  optimal = optimize(lrDifferenceAbs, 
                     maximum = FALSE, interval = c(0, max(Aseq)),
                     tol = .Machine$double.eps^0.5) 
  print(optimal)
  optimalAlphaBig = optimal$ minimum
  whichOptimal = ( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
  optimalAlphaSmall = 2*alpha - optimalAlphaBig
  cutoffsBig =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars())))
  cutoffsSmall = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars())))
  # testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
  #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
  conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]), 
                                 pow(cutoffsBig, M.alt=1, V=vars()[2]) )
  
  margPowerEqualLR = mean(conditionalPowerEqualLR)
  
  cat('results.EqualLR.Table\n')
  lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    
  results.EqualLR.Table = data.frame(
    alphas = c(optimalAlphaSmall, optimalAlphaBig), 
    cutoffs = c(cutoffsSmall, cutoffsBig), 
    LR = c(lrSmall, lrBig),
    power = conditionalPowerEqualLR, 
    marginalPower = rep(margPowerEqualLR, 2)
  ) 
  return(results.EqualLR.Table)
})
```

```{r}
output$results.EqualLR = renderTable({
  df = results.EqualLR.Table()
  names(df)[names(df) == 'marginalPower'] = 'marginal<br>power'
  names(df)[names(df) == 'power'] = 'conditional<br>power'
  df
}, sanitize.colnames.function = function(x) x)
```

```{r}
cutoffs.equalLR = reactive(results.EqualLR.Table()$cutoffs)
```


```{r}
output$lrDifferenceAbsPlot  = renderPlot({
    Aseq<- seq(0.001, 0.05, length=1000)
    plot(Aseq,
       lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
#  abline(v=alpha/2, lty=2)
  #abline(v=results.EqualLR.Table()$alphas[2], lty=3)
})
```

```{r}
# not a helpful plot

### lrDifferenceAbsPlot

#plotOutput(outputId = 'lrDifferenceAbsPlot')  #, width = '40%', height='40%')

```
<!-- ## Comparing power, as alpha allocation changes -->

```{r}
minAlphaBig = 0.002

### This plot shows how the power and cutoffs change as the alpha allocation changes
# output$theplotChangingAlphaBig = renderPlot({
#   cutoffsBig =   qnorm(1 - input$ChangingAlphaBig, sd = sqrt(max(vars())))
#   cutoffsSmall = qnorm(1 - (alpha - input$ChangingAlphaBig), sd = sqrt(min(vars())))
#   cutoffsAny = c(cutoffsSmall, cutoffsBig)
#   # testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
#   #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
#   conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]), 
#                                  pow(cutoffsBig, M.alt=1, V=vars()[2]) )
#   upperX = qnorm(1 - minAlphaBig, sd = sqrt(max(vars())))
#   cat('upperX ', upperX, '\n')
#   Cseq <- seq(lowerC, max(results.EqualLR.Table()$cutoffs), length=1000)
# 
#   plot(Cseq, ylim=c(0,1), xlim=c(0, upperX * 1.005),
#        pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
#        pch='', col=col.v1)
#   #  Draw points for the equal alpha case.
#   pow.M.alt.1 = pow(cutoffs.equalAlpha(), M.alt=1, V=vars())
#   points(cutoffs.equalAlpha(), pow.M.alt.1,  pch='A')
#   points(cutoffsAny, pow(cutoffsAny, M.alt=1, V=vars()), pch='P')
#   points(cutoffs.equalLR(), pow(cutoffs.equalLR(), M.alt=1, V=vars()), col=col.v[4], pch='L')
#   #   lines(x=c(lowerC, cutoffs()[1]), y= c(alpha,alpha) , col='black')
#   # lines(x=c(lowerC, cutoffs()[2]), y= c(alpha,alpha) , col='black')
#   # text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
#   # points(cutoffs(), c(alpha, alpha), col=col.v, pch='C')
#   #points(cutoffs, pow(cutoffs, M.alt=0), col=col.v, pch='C')  # null hypothesis
#   # points(cutoffs(), pow.M.alt.1(), col=col.v, pch='P')
#   cutoffsLR = results.EqualLR.Table()$cutoffs
#   cutoffsLRSmall =   cutoffsLR[1]
#   cutoffsLRBig =   cutoffsLR[2]
#   for(Vn in 1:2) {
#     # lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
#     # lines(x=c(lowerC, cutoffsLR[Vn]), y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
#     # lines(x=rep(cutoffs[Vn],2), y = c(alpha, pow.M.alt.1() [Vn]), col=col.v[Vn])
#     # text(x=lowerC, y=pow.M.alt.1() [Vn], col=col.v[Vn],
#     #      labels = paste('V=', vars()[Vn], '  power=', 
#     #                     signif(digits = 2, pow.M.alt.1() [Vn])), adj = c(0,0))
#     
#   }
#   ### average
#   margPowerEqualAlpha = mean(pow.M.alt.1)
#   #  lines(x=c(lowerC, mean(cutoffs)), y= rep(margPowerEqualAlpha, 2), col=col.v[3])
#   #lines(rep(lowerC, 2), pow.M.alt.1, col=col.v[3])
#   text(x=lowerC, y=mean(pow.M.alt.1), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
#        labels = paste('marginal power equal alpha:',
#                       signif(digits = 3, margPowerEqualAlpha)))
# 
#   conditionalPowerEqualLR = results.EqualLR.Table()$power
#   margPowerEqualLR = mean(conditionalPowerEqualLR)
#   points( cex=2, c(  cutoffsLRSmall,   cutoffsLRBig), 
#           conditionalPowerEqualLR, col=col.v['LR'], pch='L')
#   points( cex=3, c(  cutoffsLRSmall,   cutoffsLRBig), 
#           conditionalPowerEqualLR, col=col.v['LR'], pch=1)
#   text(x=lowerC, y=margPowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'], xpd=NA,
#        labels = paste('marginal power equal LR:',
#                       signif(digits = 3, margPowerEqualLR)))
#   legend(x='topright', 
#          legend=c(paste('Variance=', vars()[1]), 
#                   paste('Variance=', vars()[2]), 
#                   'equal alpha',
#                   'equal LR'
#          ), 
#          text.col =  col.v
#   )
# })
```

```{r echo=FALSE}
## We are FOR NOW keeping Pr(variance = vars[1] equal to 1/2).
# numericInput('ChangingAlphaBig', 'Changing Alpha allocation', value = alpha,
#              min=minAlphaBig, max=alpha * 2 - minAlphaBig, step=minAlphaBig)
# 
# plotOutput('theplotChangingAlphaBig')
```


```{r}
div(
  numericInput('varsBig', 'variance Big', value = vars.init[2], min = vars.init[1], step=vars.step),
  numericInput('varsSmall', 'variance Small', value = vars.init[1], min = 0,
               max = vars.init[2], step=vars.step)
)
vars = reactive({c(input$varsSmall, input$varsBig)})
observeEvent(input$varsBig,
             updateNumericInput(inputId='varsSmall', max = input$varsBig))
observeEvent(input$varsSmall,
             updateNumericInput(inputId='varsBig', min = input$varsSmall))
```

```{r echo=FALSE}

#### Plot power and cutoffs for equalAlpha and equalLR cases.

output$theplotForEquals = renderPlot({
    Cseq <- seq(lowerC, max(results.EqualLR.Table()$cutoffs), length=1000)

    cutoffsLR = results.EqualLR.Table()$cutoffs
    cutoffsLRSmall =   cutoffsLR[1]
    cutoffsLRBig =   cutoffsLR[2]
    cutoffsAlpha = results.EqualAlpha.Table()$cutoffs
    cutoffsAlphaSmall =   cutoffsAlpha[1]
    cutoffsAlphaBig =   cutoffsAlpha[2]
  plot(Cseq, 
       ylim=c(0,1), xlim=c(0,   cutoffsLRBig * 1.01),
       pow(Cseq, M.alt=0, 1), 
       xlab='critical rejection cutoff, conditional on which experiment',
       ylab='power to reject alternative',
       pch='', col=col.v1)
  lines(Cseq, pow(Cseq, M.alt=1, vars()[1]), col=col.v1)
  lines(Cseq, pow(Cseq, M.alt=1, vars()[2]), col=col.v2)
  #####  LR ########
  # lines(x=c(lowerC,   cutoffsLR[1]), y= c(alpha,alpha) , col='black')
  # lines(x=c(lowerC,   cutoffsLR[2]), y= c(alpha,alpha) , col='black')
  # text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  
    conditionalPowerEqualLR = results.EqualLR.Table()$power

  for(v in 1:2) {
    lines(rep(cutoffsAlpha[v],2), 
          c(alpha, pow(cutoffsAlpha[v], M.alt=0, V=vars()[v])))
    points(cutoffsAlpha[v], 
           results.EqualAlpha.Table()$power[v], col=col.v[v], pch='A', cex=2)
    text(x= cutoffsAlpha[v] + ifelse(v==1, -0.8, -0.8) ,
         y=results.EqualAlpha.Table()$power[v],
          xpd=NA,  
         cex=1, 
         'conditional power\nequal Alpha', col=col.v[v])
    text(pos=ifelse(v==1, 3, 2), xpd=NA,  cex=1, 
         cutoffsLR[v] - 0.05*(v==2), results.EqualLR.Table()$power[v]
         + 0.05*(v==1) - 0.02*(v==2), 
          'conditional power\nequal LR', col=col.v[v])
  }

    points( cex=2, c(cutoffsLRSmall, cutoffsLRBig), 
          conditionalPowerEqualLR, col=col.v, pch='L')
  
  
  # for(Vn in 1:2) {
  #   lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
  #   lines(x=c(lowerC, cutoffs.equalAlpha[Vn]), 
  #         y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
  #   lines(x=rep(cutoffs.equalAlpha[Vn],2), 
  #         y = c(alpha, pow.M.alt.1 [Vn]), col=col.v[Vn])
  #   text(x=lowerC, y=pow.M.alt.1 [Vn], col=col.v[Vn],
  #        labels = paste('V=', vars()[Vn], '  power=', 
  #                       signif(digits = 2, pow.M.alt.1 [Vn])), adj = c(0,0))
  #   
  # }
  ### average
  margPowerEqualAlpha = signif(digits=4, mean(results.EqualAlpha.Table()$power))
  margPowerEqualLR = signif( digits=4, mean(results.EqualLR.Table()$power))
  
  text(x=2.8, y = margPowerEqualLR + 0.07, adj = c(0,-1), 'marginal power:')
  text(x=2.8, y = margPowerEqualAlpha, xpd=NA, adj = c(0,1), 
       col=col.v['alpha'], xpd=NA,
       labels = paste('equal alpha:', 
                      signif(digits = 3, margPowerEqualAlpha)))
  lines(c(0, 2.8), rep(margPowerEqualAlpha,2),  col=col.v['alpha'])
  text(x=2.8, y=margPowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'],
       labels = paste('equal LR:', 
                      signif(digits = 3, margPowerEqualLR)))
  lines(c(0, 2.8), rep(margPowerEqualLR,2),  col=col.v['LR'])

  legend(x='topright', 
         legend=c(paste('Variance=', vars()[1]), 
                  paste('Variance=', vars()[2]) 
         ), 
         text.col =  c(col.v1, col.v2)
  )

})
```

## Comparing power, for equal alpha and equal LR


```{r}
centerIt = function(S) span(style='text-align:center;  display: flex;
  align-items: center;
  justify-content: center; font-size:larger', em(S))
output$meanAlphaPower = renderText(
  paste('Marginal power = ' , signif( mean(results.EqualAlpha.Table()$power))))
output$meanLRPower = renderText(
  paste('Marginal power = ', signif(mean(results.EqualLR.Table()$power))))
fluidRow(column(5, 
                centerIt('Equal alpha'), 
                 centerIt(textOutput('meanAlphaPower')), 
                tableOutput('results.EqualAlpha'),
                centerIt('Equal LR'), 
                centerIt(textOutput('meanLRPower')),
                tableOutput('results.EqualLR')
                ),
         column(6, plotOutput('theplotForEquals')))

```

So for negligible gain,  you do stupid things.
You are violating conditionality principle--
condition only on the experiment you actually did.

So what happens if the small variance inaccurate instrument is picked, and the observation X is exactly at the cutoff for same-likelihood-ratio plan with alpha = 0.05.  Well then you can claim that the P value is 0.05. But being honest, X is at the cutoff for alpha = 0.09. To paraphrase,

"The coin flip could have picked a worse experiment, and even if it didn't, the existence of the possibility means the strength of evidence is 0.05 not 0.09."


