---
title: "neyman pearson funky demo"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
options(error=print(traceback()))

tO = function(V) textOutput(inline=TRUE, V)
```


#  neyman pearson funky demo
# violating conditionality principle



```{r echo=FALSE}

h0 = 0 # null
ha = 1 #  alternative hypothesis
vars.init = c(0.4, 3) ####    Small, and large. so power is large, Small.
div(
  numericInput('varsBig', 'varsBig', value = vars.init[2], min = 0, step=0.1),
  numericInput('varsSmall', 'varsSmall', value = vars.init[1], min = 0, step=0.1)
)
vars = reactive({c(input$varsSmall, input$varsBig)})

alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
lowerC = 0
Cseq <- seq(lowerC,3,length=1000)

likelihood = function(h, v, x) {
  1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
# likelihood(h = 0, v = 1, x = qnorm(1-0.025)) # yes 0.025.
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
```

```{r eval=FALSE}

### skip this plot
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall())))
# abline(v=0:1)
# lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsBig() )))
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall() )))

```

The setting:  one observation from a normal distribution, mean unknown.
If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
```{r echo=FALSE}
output[['varsSmall.text']] = renderText(input[['varsSmall']])   # OK.
output[['varsSmall.text.2']] = renderText(input[['varsSmall']])   # OK.

output[['varsBig.text']] = renderText(input[['varsBig']])   # OK.
output[['varsBig.text.2']] = renderText(input[['varsBig']])   # OK.

cutoffs  = reactive({
  qnorm(1 - alpha, sd = sqrt(vars()))
})

output$cutoffs.text = renderText(paste0(signif(digits=3, cutoffs()[1]), 
                                        ' and ',
                                   signif(digits=3, cutoffs()[2] ) ) )


output$results.EqualAlpha = renderTable({
  #cutoffs  = qnorm(1 - alpha, sd = sqrt(vars))
  lrEqualAlpha <<-  lr(cutoffs(), vars())
  pow.M.alt.1 <<- reactive(pow(cutoffs(), M.alt=1, V=vars()) )
  results.EqualAlpha <<- data.frame(
    alphas = rep(alpha, 2),
    cutoffs=cutoffs(), 
    LR= lrEqualAlpha, 
    power=pow.M.alt.1(),
    marginalPower= rep(mean(pow.M.alt.1()), 2))
  results.EqualAlpha
})

```
But we have **two** instruments to measure *X*, with different variances
**`r textOutput(inline=T, 'varsSmall.text')` and `r textOutput(inline=T, 'varsBig.text')`.**  
They could be different in precisions. Or they could be because there are two different possible sample sizes.  *The instrument used is picked at random!*  ---- **But we know which one was picked!**

Suppose we use the same *alpha*, the same conditional type I error rate, regardless of which instrument. If you know which instrument was picked, you would of course use the relevant significance cutoff. Here for *alpha* = `r alpha`, used regardless of the instrument chosen, these two conditional cutoffs are `r textOutput(inline=T, 'cutoffs.text')`.  
Then regardless of the chances for the two instrument, the overall marginal *alpha* is the same.
```{r}
output$lrEqualAlpha.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha$LR[1]),
         ' and ',
         signif(digits=3, results.EqualAlpha$LR[2])
         )
})
```

The likelihood ratios at the two critical values are `r textOutput(inline=TRUE, 'lrEqualAlpha.text')`.
Not the same.  Should we maximize power by  forcing the two likelihood ratios to be the same?

If we raise the cutoff for the Bigger variance `r textOutput(inline=T, 'varsBig.text.2')`, and lower it for the Smaller variance `r textOutput(inline=T, 'varsSmall.text.2')`, watch what happens. The likelihood ratios get closer and closer.

```{r}
  lrDifference = function(alphaBig){
      lrBigValues = numeric(0)
      lrSmallValues = numeric(0)
      cutoffsBigValues = numeric(0)
      cutoffsSmallValues = numeric(0)
      
    alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
    cutoffsBig = qnorm(1 - alphaBig, sd = sqrt(max(vars())))
    cutoffsSmall = qnorm(1 - alphaSmall, sd = sqrt(min(vars())))
    cutoffsBigValues =  append(cutoffsBigValues, cutoffsBig)
    cutoffsSmallValues =  append(cutoffsSmallValues, cutoffsSmall)
    #print(c( lr(cutoffsBig, v = max(vars)) , lr(cutoffsSmall, v = min(vars)) ) )
    lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    lrBigValues =  append(lrBigValues, lrBig)
    lrSmallValues =  append(lrSmallValues, lrSmall)
    lrBig - lrSmall
  }
  lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
```
```{r}
  results.EqualLR.Table = reactive({
  Aseq<- seq(0.001, 0.05, length=1000)
  optimal = optimize(lrDifferenceAbs, 
                     maximum = FALSE, interval = c(0, max(Aseq)),
                     tol = .Machine$double.eps^0.5) 
  print(optimal)
  optimalAlphaBig = optimal$ minimum
  whichOptimal = ( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
  optimalAlphaSmall = 2*alpha - optimalAlphaBig
  cutoffsBig =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars())))
  cutoffsSmall = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars())))
  # testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
  #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
  conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]), 
                                 pow(cutoffsBig, M.alt=1, V=vars()[2]) )
  
  averagePowerEqualLR = mean(conditionalPowerEqualLR)
  
  cat('results.EqualLR.Table\n')
  lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    
  print(data.frame(
    alphas = c(optimalAlphaSmall, optimalAlphaBig), 
    cutoffs = c(cutoffsSmall, cutoffsBig), 
    LR = c(lrSmall, lrBig),
    power = conditionalPowerEqualLR, 
    marginalPower = rep(averagePowerEqualLR, 2)
  ) )
})
```

```{r}
output$results.EqualLR = renderTable({
  results.EqualLR.Table()
})
```



```{r}
output$lrDifferenceAbsPlot  = renderPlot({
    Aseq<- seq(0.001, 0.05, length=1000)
    plot(Aseq,
       lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
#  abline(v=alpha/2, lty=2)
  #abline(v=results.EqualLR.Table()$alphas[2], lty=3)
})
```

```{r}
# not a helpful plot

### lrDifferenceAbsPlot

#plotOutput(outputId = 'lrDifferenceAbsPlot')  #, width = '40%', height='40%')

```
## Comparing power, as alpha allocation changes

```{r}
minAlphaBig = 0.002

output$theplotChangingAlphaBig = renderPlot({
  cutoffsBig =   qnorm(1 - input$ChangingAlphaBig, sd = sqrt(max(vars())))
  cutoffsSmall = qnorm(1 - (alpha - input$ChangingAlphaBig), sd = sqrt(min(vars())))
  cutoffs = c(cutoffsSmall, cutoffsBig)
  # testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
  #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
  conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]), 
                                 pow(cutoffsBig, M.alt=1, V=vars()[2]) )
  upperX = qnorm(1 - minAlphaBig, sd = sqrt(max(vars())))
  plot(Cseq, ylim=c(0,1), xlim=c(0, upperX * 1.005),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
  
  #   lines(x=c(lowerC, cutoffs()[1]), y= c(alpha,alpha) , col='black')
  # lines(x=c(lowerC, cutoffs()[2]), y= c(alpha,alpha) , col='black')
  # text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  # points(cutoffs(), c(alpha, alpha), col=col.v, pch='C')
  #points(cutoffs, pow(cutoffs, M.alt=0), col=col.v, pch='C')
  # points(cutoffs(), pow.M.alt.1(), col=col.v, pch='P')
  
  
  cutoffsLR = results.EqualLR.Table()$cutoffs
  cutoffsLRSmall =   cutoffsLR[1]
  cutoffsLRBig =   cutoffsLR[2]
  for(Vn in 1:2) {
    lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
    lines(x=c(lowerC, cutoffs[Vn]), y= rep(pow.M.alt.1() [Vn], 2), col=col.v[Vn])
    # lines(x=rep(cutoffs[Vn],2), y = c(alpha, pow.M.alt.1() [Vn]), col=col.v[Vn])
    text(x=lowerC, y=pow.M.alt.1() [Vn], col=col.v[Vn],
         labels = paste('V=', vars()[Vn], '  power=', 
                        signif(digits = 2, pow.M.alt.1() [Vn])), adj = c(0,0))
    
  }

  ### average
  averagePowerEqualAlpha = mean(pow.M.alt.1())
  #  lines(x=c(lowerC, mean(cutoffs)), y= rep(averagePowerEqualAlpha, 2), col=col.v[3])
  lines(rep(lowerC, 2), pow.M.alt.1(), col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1()), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal alpha:', 
                      signif(digits = 3, averagePowerEqualAlpha)))
  
  conditionalPowerEqualLR = results.EqualLR.Table()$power
  averagePowerEqualLR = mean(conditionalPowerEqualLR)
  points( cex=2, c(  cutoffsLRSmall,   cutoffsLRBig), 
          conditionalPowerEqualLR, col=col.v['LR'], pch='L')
  points( cex=3, c(  cutoffsLRSmall,   cutoffsLRBig), 
          conditionalPowerEqualLR, col=col.v['LR'], pch=1)
  text(x=lowerC, y=averagePowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'], xpd=NA,
       labels = paste('marginal power equal LR:', 
                      signif(digits = 3, averagePowerEqualLR)))
  # arrowTerminus = c(1.2, averagePowerEqualLR+0.1)
  # diagram::curvedarrow(to=arrowTerminus, from=c(cutoffsSmall,
  #                                            conditionalPowerEqualLR[1]) ,
  #                      lwd = 2, lty = 1, lcol = col.v['LR'], 
  #           arr.col = col.v['LR'], arr.pos = 1, curve = -.3, dr = 0.01, 
  #           endhead = FALSE, segment = c(0, 1))
  # diagram::curvedarrow(to=arrowTerminus, from=c(cutoffsBig,
  #                                            conditionalPowerEqualLR[2]) ,
  #                      lwd = 2, lty = 1, lcol = col.v['LR'], 
  #           arr.col = col.v['LR'], arr.pos = 1, curve = .1, dr = 0.01, 
  #           endhead = FALSE, segment = c(0, 1))
  # 
  legend(x='topright', 
         legend=c(paste('Variance=', vars()[1]), 
                  paste('Variance=', vars()[2]), 
                  'equal alpha',
                  'equal LR'
         ), 
         text.col =  col.v
  )
})
```

```{r echo=FALSE}
numericInput('ChangingAlphaBig', 'Changing Alpha allocation', value = alpha/2, min=minAlphaBig, max=alpha, step=minAlphaBig)

plotOutput('theplotChangingAlphaBig')
```


```{r echo=FALSE}
output$theplotForEquals = renderPlot({
    cutoffsLR = results.EqualLR.Table()$cutoffs
    cutoffsLRSmall =   cutoffsLR[1]
    cutoffsLRBig =   cutoffsLR[2]
  plot(Cseq, ylim=c(0,1), xlim=c(0,   cutoffsLRBig * 1.01),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
    lines(x=c(lowerC,   cutoffsLR[1]), y= c(alpha,alpha) , col='black')
  lines(x=c(lowerC,   cutoffsLR[2]), y= c(alpha,alpha) , col='black')
  text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  points(cutoffs(), c(alpha, alpha), col=col.v, pch='C')
  #points(cutoffs, pow(cutoffs, M.alt=0), col=col.v, pch='C')
  points(cutoffs(), pow.M.alt.1(), col=col.v, pch='P')
  
  
  for(Vn in 1:2) {
    lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
    lines(x=c(lowerC, cutoffs[Vn]), y= rep(pow.M.alt.1() [Vn], 2), col=col.v[Vn])
    lines(x=rep(cutoffs[Vn],2), y = c(alpha, pow.M.alt.1() [Vn]), col=col.v[Vn])
    text(x=lowerC, y=pow.M.alt.1() [Vn], col=col.v[Vn],
         labels = paste('V=', vars()[Vn], '  power=', 
                        signif(digits = 2, pow.M.alt.1() [Vn])), adj = c(0,0))
    
  }
  ### average
  averagePowerEqualAlpha = mean(pow.M.alt.1())
  #  lines(x=c(lowerC, mean(cutoffs)), y= rep(averagePowerEqualAlpha, 2), col=col.v[3])
  lines(rep(lowerC, 2), pow.M.alt.1(), col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1()), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal alpha:', 
                      signif(digits = 3, averagePowerEqualAlpha)))
  
  conditionalPowerEqualLR = results.EqualLR.Table()$power
  averagePowerEqualLR = mean(conditionalPowerEqualLR)
  points( cex=2, c(cutoffsSmall, cutoffsBig), 
          conditionalPowerEqualLR, col=col.v['LR'], pch='L')
  points( cex=3, c(cutoffsSmall, cutoffsBig), 
          conditionalPowerEqualLR, col=col.v['LR'], pch=1)
  text(x=lowerC, y=averagePowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'], xpd=NA,
       labels = paste('marginal power equal LR:', 
                      signif(digits = 3, averagePowerEqualLR)))
  # arrowTerminus = c(1.2, averagePowerEqualLR+0.1)
  # diagram::curvedarrow(to=arrowTerminus, from=c(cutoffsSmall,
  #                                            conditionalPowerEqualLR[1]) ,
  #                      lwd = 2, lty = 1, lcol = col.v['LR'], 
  #           arr.col = col.v['LR'], arr.pos = 1, curve = -.3, dr = 0.01, 
  #           endhead = FALSE, segment = c(0, 1))
  # diagram::curvedarrow(to=arrowTerminus, from=c(cutoffsBig,
  #                                            conditionalPowerEqualLR[2]) ,
  #                      lwd = 2, lty = 1, lcol = col.v['LR'], 
  #           arr.col = col.v['LR'], arr.pos = 1, curve = .1, dr = 0.01, 
  #           endhead = FALSE, segment = c(0, 1))
  
  legend(x='topright', 
         legend=c(paste('Variance=', vars()[1]), 
                  paste('Variance=', vars()[2]), 
                  'equal alpha',
                  'equal LR'
         ), 
         text.col =  col.v
  )
})
```

## Comparing power, for equal alpha and equal LR


```{r}
centerIt = function(S) span(style='text-align:center;  display: flex;
  align-items: center;
  justify-content: center; font-size:larger', em(S))
fluidRow(column(5, centerIt('Equal alpha'), tableOutput('results.EqualAlpha'),
         centerIt('Equal LR'), tableOutput('results.EqualLR')),
         column(6, plotOutput('theplotForEquals')))

```


