---
title: "Untitled"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


#  neyman pearson funky demo
# violating conditionality principle



```{r echo=FALSE}

h0 = 0  # null
ha = 1 #  alternative hypothesis
vars = c(0.4, 3) ####    Small, and large. so power is large, Small.
alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
  lowerC = 0
  Cseq <- seq(lowerC,3,length=1000)

likelihood = function(h, v, x) {
   1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
# likelihood(h = 0, v = 1, x = qnorm(1-0.025)) # yes 0.025.
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
```

```{r eval=FALSE}

### skip this plot
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))
abline(v=0:1)
lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[2])))
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))

```

The setting:  one observation from a normal distribution, mean unknown.
If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
```{r echo=FALSE}
  C95  = qnorm(1 - alpha, sd = sqrt(vars))
lrEqualAlpha =  lr(C95, vars)
  pow.M.alt.1 = pow(C95, M.alt=1, V=vars)

```
But we have **two** instruments to measure *X*, with different variances  `r vars[1]` and `r vars[2]`.  They could be different in precisions. Or they could be because there are two different possible sample sizes.  *The instrument used is picked at random!*  ---- **But we know which one was picked!**

Suppose we use the same *alpha*, the same conditional type I error rate, regardless of which instrument. If you know which instrument was picked, you would of course use the relevant significance cutoff. Here for *alpha* = `r alpha`, used regardless of the instrument chosen, these two conditional cutoffs are `r signif(digits=4, C95[1])` and `r signif(digits=4, C95[2])`.  Then regardless of the chances for the two instrument, the overall marginal *alpha* is the same.

The likelihood ratios at the two critical values are `r lrEqualAlpha[1]` and `r lrEqualAlpha[2]`.  Not the same.  Should we maximize power by  forcing the two likelihood ratios to be the same?

If we raise the cutoff for the Bigger variance, and lower it for the Smaller variance, watch what happens. The likelihood ratios get closer and closer.

```{r}
Aseq<- seq(0.001, 0.05, length=1000)
lrBigValues = numeric(0)
lrSmallValues = numeric(0)
C95BigValues = numeric(0)
C95SmallValues = numeric(0)
lrDifference = function(alphaBig){
  alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
  C95Big = qnorm(1 - alphaBig, sd = sqrt(max(vars)))
  C95Small = qnorm(1 - alphaSmall, sd = sqrt(min(vars)))
  C95BigValues =  append(C95BigValues, C95Big)
  C95SmallValues =  append(C95SmallValues, C95Small)
  #print(c( lr(C95Big, v = max(vars)) , lr(C95Small, v = min(vars)) ) )
  lrBig = lr(C95Big, v = max(vars)) 
  lrSmall = lr(C95Small, v = min(vars))
  lrBigValues =  append(lrBigValues, lrBig)
  lrSmallValues =  append(lrSmallValues, lrSmall)
  lrBig - lrSmall
}
lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
plot(Aseq,
     lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
optimalAlphaBig = optimize(lrDifferenceAbs, 
                           maximum = FALSE, interval = c(0, max(Aseq))) $ minimum
whichOptimal = print( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
optimalAlphaSmall = 2*alpha - optimalAlphaBig
C95Big =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars)))
C95Small = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars)))
# testthat::expect_equal(lr(C95Small, min(vars)), 
#                        lr(C95Big, max(vars)) , tolerance = 1e-3)
conditionalPowerEqualLR =   c( pow(C95Small, M.alt=1, V=vars[1]), 
                               pow(C95Big, M.alt=1, V=vars[2]) )

averagePowerEqualLR = mean(conditionalPowerEqualLR)
  
results = list(C95Small, C95Big, 
               conditionalPowerEqualLR, averagePowerEqualLR)
```

The plot shows the absolute difference in the two likelihood ratios, as the proportion of **alpha** for the big variance. We see that the likelihood ratios are equal when the alpha for the big variance instrument is **`r round(optimalAlphaSmall, digits=3)`**. Then the other **alpha**, for the less precise instrument, is `r round(optimalAlphaBig, digits=3)`. They average to **alpha = `r alpha`**.

```{r echo=FALSE}
### conditional power

output$theplot = renderPlot({
  plot(Cseq, ylim=c(0,1), xlim=c(0, C95Big * 1.01),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
  lines(x=c(lowerC, C95[1]), y= c(alpha,alpha) , col='black')
  lines(x=c(lowerC, C95[2]), y= c(alpha,alpha) , col='black')
  text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  points(C95, c(alpha, alpha), col=col.v, pch='C')
  #points(C95, pow(C95, M.alt=0), col=col.v, pch='C')
  points(C95, pow.M.alt.1, col=col.v, pch='P')
  
  
  for(Vn in 1:2) {
    lines(Cseq, pow(Cseq, M.alt=1, V=vars[Vn]), lty=2, lwd=3, col=col.v[Vn])
    lines(x=c(lowerC, C95[Vn]), y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
    lines(x=rep(C95[Vn],2), y = c(alpha, pow.M.alt.1 [Vn]), col=col.v[Vn])
    text(x=lowerC, y=pow.M.alt.1 [Vn], col=col.v[Vn],
         labels = paste('V=', vars[Vn], '  power=', 
                        signif(digits = 2, pow.M.alt.1 [Vn])), adj = c(0,0))
    
  }
  ### average
  averagePowerEqualAlpha = mean(pow.M.alt.1)
#  lines(x=c(lowerC, mean(C95)), y= rep(averagePowerEqualAlpha, 2), col=col.v[3])
  lines(rep(lowerC, 2), pow.M.alt.1, col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal alpha:', 
                      signif(digits = 6, averagePowerEqualAlpha)))

  
  points( cex=2, c(C95Small, C95Big), 
          conditionalPowerEqualLR, col=col.v[4], pch='L')
  text(x=lowerC, y=averagePowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v[4], xpd=NA,
       labels = paste('marginal power equal LR:', 
                      signif(digits = 6, averagePowerEqualLR)))
  
  
  legend(x='topright', 
         legend=c(paste('Variance=', vars[1]), 
                  paste('Variance=', vars[2]), 
                  'equal alpha',
                  'equal LR'
         ), 
         text.col =  col.v
  )
})

output$likrat = renderPlot({
  plot(c(C95SmallValues, C95BigValues), 
       )
})
plotOutput('theplot')
```



If the variance is `r vars[1]`, the cutoff for alpha=0.05 is `r C95[1] `.  
But if the variance is `r vars[2]`, the cutoff for alpha=0.05 is `r C95[2] `.  
The conditional statistical **powers** for rejecting the "null" (bland) hypothesis $M=0$  are `r pow.M.alt.1[1]` and `r pow.M.alt.1[2]`,


and the average power, averaging the conditional powers
over  which variance is chosen (50-50), is 

## Likelihood ratios
The likelihood ratios are, respectively, 

`r lr(C95[1], vars[1])` and `r lr(C95[2], vars[2])`

Not equal!!

The equalizer for the likelihood ratio is when  
AlphaBig = optimalAlphaBig = `r optimalAlphaBig` and  
AlphaSmall = optimalAlphaSmall = `r optimalAlphaSmall`.  
The critical value for the small variance is now `r C95Small`,
and for the large variance it's `r C95Big`.


Are you content to report a P value based on the marginal **alpha** when you KNOW which instrument was used? So the **conditionality principle** is used FIRST.



