#  neyman pearson funky demo
# violating conditionality principle



```{r}
vars = c(1, 2)  ### two different variances
h0 = 0  # null
ha = 1 #  alternative hypothesis
likelihood = function(h, v, x) {
   1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
likelihood(h = 0, v = 1, x = qnorm(1-0.025))
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))
abline(v=0:1)
lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[2])))
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))

```

If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$


```{r}
### conditional power
col.v1 = 'red'
col.v2 = 'green'
col.v = c(col.v1, col.v2)
pow = function(C, M.alt=1, V=vars[1]) {
  1 - pnorm(C,  M.alt,sqrt(V)) 
}
lowerC = 0
Cseq <- seq(lowerC,3,length=1000)
plot(Cseq, 
     pow(Cseq, M.alt=0), xlab='critical rejection cutoff', ylab='power',
     pch='', col=col.v1)
C95  = qnorm(1 - alpha, sd = sqrt(vars))
lines(x=c(lowerC, C95[1]), y= c(alpha,alpha) , col='black')
lines(x=c(lowerC, C95[2]), y= c(alpha,alpha) , col='black')
text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  points(C95, c(alpha, alpha), col=col.v, pch='C')
  #points(C95, pow(C95, M.alt=0), col=col.v, pch='C')
pow.M.alt.1 = pow(C95, M.alt=1, V=vars)
points(C95, pow.1.1, col='black', pch='C')

  
for(Vn in 1:2) {
  lines(Cseq, pow(Cseq, M.alt=1, V=vars[Vn]), lty=2, lwd=3, col=col.v[Vn])
}


#abline(h=0.5,v=1, lty=2, col='lightgrey')
#  lines(Cseq, pow(Cseq, M.alt=1), lty=2, col=col.v1)
  lines(Cseq, pow(Cseq, M.alt=1, V=vars[Vn]), lty=2, lwd=3, col=col.v[Vn])

  
  points(C95, pow.1.1, col='black', pch='C')
  lines(x=c(lowerC, C95), y= rep(pow.1.1, 2) , col='black')
  text(x=lowerC, y=pow.1.1, labels = paste('power=', signif(digits = 2, pow.1.1)), adj = c(0,0), col='black') 
#  pow.1.2 = pow(C95, M.alt=1, V=vars[2])
  points(C95.2, pow.1.2, col='green', pch='C')
  lines(x=c(lowerC, C95), y= rep(pow.1.2, 2) , col='green')
  text(x=lowerC, y=pow.1.2, labels = paste('power=', signif(digits = 2, pow.1.2)), adj = c(0,0), col='green') 
}
  legend(x='topright', 
         legend=c('M.alt=1 V=1', 
                  'M.alt=1 V=2', 
                  'M.alt=0 V=1' , 
                  'M.alt=0 V=2' ), 
         text.col =  c(col.v1, col.v2, col.v1, col.v2)
  )

### unconditional power
```

```{r}
cutoffs = qnorm(0.95, 0, sqrt(vars))
```

If V=1, the cutoff for alpha=0.05 is `r cutoffs[1] `.
But if V=2, the cutoff for alpha=0.05 is `r cutoffs[2] `.
The likelihood ratios are, respectively, 

`r lr(cutoffs[1], vars[1])` and `r lr(cutoffs[2], vars[2])`

The average power, conditioning on which variance is chosen:
$$
Power()
$$
Now lets


