---
title: "The beautiful Neyman Pearson lemma <br>makes you do nutty things"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
resource_files:
- www/KeyHandler.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(DT)  # not data.table
library(reactable)
library(shinyDebuggingPanel, quietly = TRUE)
library(keys)

options(error=print(traceback()))
options(shiny.launch.browser=TRUE)


tO = function(V) textOutput(inline=TRUE, V)
```


```{r keys, echo=FALSE}
### none of these works -- initially.
shiny::includeScript(path =  'www/KeyHandler.js')  
    # works at shinyapps.io
    ### and locally, all of a sudden!
## Did I have to deploy first before running this line locally would work?
```



```{r shinyDebuggingPanel}
observeEvent(input$ctrlDpressed, {}) # just to flush the ctrl-D press.

shinyDebuggingPanel::makeDebuggingPanelOutput(
   session, toolsInitialState = TRUE,
   condition='ctrlDpressed === true')

fluidRow(
shinyDebuggingPanel::withDebuggingPanel()
)



```

```{r likelihoods, echo=FALSE}
h0 = 0 # bland, null hypothesis
ha = 1 #  exciting, alternative hypothesis
vars.init = c(small=0.4, big=3) ####    Small, and large. so power is large, Small.
vars.step = 0.1

vars = reactive({c(input$varsSmall, input$varsBig)})
observeEvent(input$varsBig,
             updateNumericInput(inputId='varsSmall', max = input$varsBig))
observeEvent(input$varsSmall,
             updateNumericInput(inputId='varsBig', min = input$varsSmall))

alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
lowerC = 0

likelihood = function(h, v, x) {
  dnorm(x = x, mean = h, sd = sqrt(v))
}
# likelihood(h = 0, v = 1, x = qnorm(1-0.025)) # yes 0.025.
lr = function(x, v) likelihood(h = ha, v=v , x=x ) / likelihood(h = h0, v=v, x=x )
# plot(Cseq <- seq(0, 1.5, length=1000), lr(Cseq, 1) )    makes sense
# lines(Cseq <- seq(0, 1.5, length=1000), lr(Cseq, v = 2) , col='red')
# lr(x = .5, v = 1)   Equals 1, correctly.
```

```{r eval=FALSE}

### skip this plot
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall())))
# abline(v=0:1)
# lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsBig() )))
# plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, varsSmall() )))

```

### The setting:  
One observation is made from a normal distribution, with mean unknown.  
Hypotheses are:  

* Null (boring): &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;mean = M = 0.  
* Alternative (exciting):   mean = M = $M_{alt}$ = 1.

What's the 'most powerful test'?


If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For the  critical value $C$ setting the rejection region, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
```{r results.EqualAlpha.Table, echo=FALSE}
output[['varsSmall.text']] = renderText(input[['varsSmall']])   # OK.
output[['varsSmall.text.2']] = renderText(input[['varsSmall']])   # OK.

output[['varsBig.text']] = renderText(input[['varsBig']])   # OK.
output[['varsBig.text.2']] = renderText(input[['varsBig']])   # OK.

cutoffs.by.alpha = function(alphaBoth, variancesBoth)
    qnorm(1 - alphaBoth, sd = sqrt(variancesBoth))


cutoffs.equalAlpha = reactive({
  cutoffs.by.alpha(c(alpha, alpha), vars())
})
cutoffs.equalAlpha_1 = reactive({
  cutoffs.by.alpha(alpha, vars()[1])
})
cutoffs.equalAlpha_2 = reactive({
  cutoffs.by.alpha(alpha, vars()[2])
})


output$cutoffsAlpha.text = renderText(paste0(signif(digits=3,
                                                    cutoffs.equalAlpha()[1]), 
                                             ' and ',
                                             signif(digits=3, cutoffs.equalAlpha()[2] ) ) )
output$cutoffsAlpha1.text = renderText(paste0(signif(digits=3, cutoffs.equalAlpha_1)))
output$cutoffsAlpha2.text = renderText(paste0(signif(digits=3, cutoffs.equalAlpha_2)))

results.EqualAlpha.Table = reactive({
    #cutoffs  = qnorm(1 - alpha, sd = sqrt(vars))
  lrEqualAlpha <<-  lr(cutoffs.equalAlpha(), vars())
  pow.M.alt.1 <<- pow(cutoffs.equalAlpha(), M.alt=1, V=vars())
  results.EqualAlpha <<- #data.frame(
    #data = 
#      signif(digits=2, 
    data.frame(
        alphas = rep(alpha, 2),
        cutoffs=cutoffs.equalAlpha(), 
        LR= lrEqualAlpha, 
        power=pow.M.alt.1,
        marginalPower= rep(mean(pow.M.alt.1), 2)
      )
    for(i in ncol(results.EqualAlpha))
      results.EqualAlpha[[i]] = signif( digits=3, results.EqualAlpha[[i]])
  # newNames = colDef(names(results.EqualAlpha))
  # print(newNames)
  # newNames[[5]] = colDef('marginal<br>power', html=TRUE)
  # print(newNames)
  # names(newNames = names(results.EqualAlpha))
  # print(newNames)
  (results.EqualAlpha )
            #, columns = newNames )
  })

output$results.EqualAlpha = renderTable({
  df = results.EqualAlpha.Table()
  names(df)[names(df) == 'marginalPower'] = 'marginal<br>power'
  names(df)[names(df) == 'power'] = 'conditional<br>power'
  df
}, sanitize.colnames.function = function(x) x)


```

### Complication!

But we have **two** instruments to measure *X*, with different variances
**`r textOutput(inline=T, 'varsSmall.text')` and `r textOutput(inline=T, 'varsBig.text')`.**  
They could be different in precisions. Or they could be because there are two different possible sample sizes.  

Even a clinical trial where the "mean" we want to know is, say, a log odds ratio, and the normal approximation is good enough.  
*A statistician and a clinical researcher walk into a bar.  *  
They're planning a clinical trial.
How many patients do they need?
**Big argument.**  The statistician wants a nice big 1000 patients, 500 treated 500 placebo.  The researcher says no way, 100 should be fine, I'm still paying off med school.  

So they flip a coin to decide.


So what's the problem?

*The instrument (or sample size) used is picked at random!*  

---- **And...  we know which one was picked!**

Suppose we use the same *alpha*, the same conditional type I error rate, regardless of which instrument was selected. We'll call that analysis plan
**"equal alpha"**.  
If you know which instrument was picked, you will of course use the relevant significance cutoff.  

#### *Right?*  

Here for *alpha* = Type I error = power to reject the bland hypothesis = `r alpha`, used regardless of the instrument chosen, the overall **marginal** *alpha* is the same.  Even if the chances for selecting either instrument are not 50-50, still the average of *alpha* with *alpha* is ... *alpha*.
```{r lrEqualAlpha.text}
# lrEqualAlpha <<-  lr(cutoffsAlpha(), vars())

# pow.M.alt.1 <<- reactive(pow(cutoffsAlpha(), M.alt=1, V=vars()) )

output$lrEqualAlpha.text = renderUI({
  span(
    span(style=paste('color:', col.v['big'], ';' ),
         signif(digits=3, results.EqualAlpha.Table()$LR[2])
         
    ),
    ' and ',
    span(style=paste('color:', col.v['small'] ),
         
         signif(digits=3, results.EqualAlpha.Table()$LR[1])
    )
  )
})
output$lrEqualAlpha_1.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha.Table()$LR[1]) )
})
output$lrEqualAlpha_2.text = renderText({
  paste0(signif(digits=3, results.EqualAlpha.Table()$LR[2]) )
})

```

Of course we want to pick the test method that maximizes the statistical power?  
Our bland null  hypothesis is  **mean = 0**.  
Pick the exciting alternative hypothesis which we want power against to be **mean = 1**.  

```{r numericinputs 1}
fluidRow(
 column(5, style=paste('color:', col.v['big']), numericInput('varsBig', 'Big variance (poor precision) ', value = vars.init['big'], min = vars.init['big'], step=vars.step) ),
 column(5, style=paste('color:', col.v['small']), 
  numericInput('varsSmall', 'Small variance (good precision)', value = vars.init['small'], min = 0,
               max = vars.init['small'], step=vars.step) )
)
```

Our current rejection region is the union of two rays going off to the right to infinity: 
```{r rejectionRegion.text}
output$rejectionRegion.text = renderUI(
  HTML(paste0(
    span(style='color:darkgreen', 
         'variance = ', vars()[2],
         " AND   X > ", 
         signif(digits=3,
                cutoffs.equalAlpha_2())
    ), 
         ' &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;OR &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;',
    span(style='color:red', 
         'variance = ', vars()[1],
         " AND   X > ", 
         signif(digits=3, cutoffs.equalAlpha_1()
         ) )
  )
))
    
div(style='text-align:center',
    uiOutput('rejectionRegion.text')
)
```


```{r plotRejectionRegion}
output$plotRejectionRegion = renderPlot( height=200, #width='100%', 
                                        {
  print(vars.init)
  plot.default( xlab= 'observed value of X',
                ylab = 'variance',
                xlim=c(0, results.EqualAlpha.Table()$cutoffs[2] * 2),
                ylim = vars.init+ c( -0.1, 0.1),
    axes=F,
    c( 0, results.EqualAlpha.Table()$cutoffs[2] * 2), 
       vars.init+ c( -0.1, 0.1), col=NA) 
#  abline(h=vars.init)  
  for (v in 1:2) {
    lines(c(0, results.EqualAlpha.Table()$cutoffs[v]), rep(vars()[v], 2),
          col = col.v[v])
    lines(c(results.EqualAlpha.Table()$cutoffs[v],
            results.EqualAlpha.Table()$cutoffs[2] * 2),  
          rep(vars()[v], each=2) , col = col.v[v],
          lwd=4)
  }
  text( x=results.EqualAlpha.Table()$cutoffs, 
        y=vars(),  pos=3,
        #adj=c(-1,-1),
        signif(digits=3, results.EqualAlpha.Table()$cutoffs),
        col=col.v[1:2], cex=1.5, xpd=NA) 
  points( results.EqualAlpha.Table()$cutoffs, 
          vars.init, col=col.v[1:2], cex=2, pch=1, xpd=NA) 
  text(results.EqualAlpha.Table()$cutoffs[2], vars()[2],
       'rejection region, var = 3.0' ,  adj=c(-1,-1), xpd=NA)
  text(results.EqualAlpha.Table()$cutoffs[1], vars()[1],
       'rejection region, var = 0.4',   adj=c(-1,-1), xpd=NA)
  text(0.05, vars()[1] - 0.3, '0', xpd=NA)  #zero
} )
plotOutput('plotRejectionRegion', height = 200)
```


The likelihood ratios at the two critical values are
`r uiOutput(inline=TRUE, 'lrEqualAlpha.text')`.   

*These numbers are NOT the same. *  

The **"Neyman-Pearson lemma"** says, you maximize power by setting the rejection region boundary to have the same likelihood ratio everywhere on the boundary.
Should we maximize power by  forcing the two likelihood ratios to be the same?

*Let's do it! *

## How does the average power change as alpha allocation changes?

We will raise the cutoff for the instrument with the Bigger variance `r textOutput(inline=T, 'varsBig.text.2')`, and lower it for the instrument with the Smaller variance `r textOutput(inline=T, 'varsSmall.text.2')`.
The average, or marginal *alpha* stays the same.  

We'll see what happens as the likelihood ratios get closer and closer.

```{r lrDifferenceAbs}
  lrDifference = function(alphaBig){
      lrBigValues = numeric(0)
      lrSmallValues = numeric(0)
      cutoffsBigValues = numeric(0)
      cutoffsSmallValues = numeric(0)
      
    alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
    cutoffsBig = qnorm(1 - alphaBig, sd = sqrt(max(vars())))
    cutoffsSmall = qnorm(1 - alphaSmall, sd = sqrt(min(vars())))
    cutoffsBigValues =  append(cutoffsBigValues, cutoffsBig)
    cutoffsSmallValues =  append(cutoffsSmallValues, cutoffsSmall)
    #print(c( lr(cutoffsBig, v = max(vars)) , lr(cutoffsSmall, v = min(vars)) ) )
    lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    lrBigValues =  append(lrBigValues, lrBig)
    lrSmallValues =  append(lrSmallValues, lrSmall)
    lrBig - lrSmall
  }
  lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
```



```{r results.EqualLR.Table}
results.EqualLR.Table = reactive({
  Aseq<- seq(0.001, 0.05, length=1000)
  optimal = optimize(lrDifferenceAbs, 
                     maximum = FALSE, interval = c(0, max(Aseq)),
                     tol = .Machine$double.eps^0.5) 
  print(optimal)
  optimalAlphaBig = optimal$ minimum
  whichOptimal = ( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
  optimalAlphaSmall = 2*alpha - optimalAlphaBig
  cutoffsBig =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars())))
  cutoffsSmall = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars())))
  # testthat::expect_equal(lr(cutoffsSmall, min(vars)), 
  #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
  conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]), 
                                 pow(cutoffsBig, M.alt=1, V=vars()[2]) )
  
  margPowerEqualLR = mean(conditionalPowerEqualLR)
  
  cat('results.EqualLR.Table\n')
  lrBig = lr(cutoffsBig, v = max(vars())) 
    lrSmall = lr(cutoffsSmall, v = min(vars()))
    
  results.EqualLR.Table = data.frame(
    alphas = c(optimalAlphaSmall, optimalAlphaBig), 
    cutoffs = c(cutoffsSmall, cutoffsBig), 
    LR = c(lrSmall, lrBig),
    power = conditionalPowerEqualLR, 
    marginalPower = rep(margPowerEqualLR, 2)
  ) 
  return(results.EqualLR.Table)
})
```

```{r results.EqualLR}
output$results.EqualLR = renderTable({
  df = results.EqualLR.Table()
  names(df)[names(df) == 'marginalPower'] = 'marginal<br>power'
  names(df)[names(df) == 'power'] = 'conditional<br>power'
  df
}, sanitize.colnames.function = function(x) x)
```

```{r cutoffs.equalLR}
cutoffs.equalLR = reactive(results.EqualLR.Table()$cutoffs)
```


```{r lrDifferenceAbsPlot render not used}
output$lrDifferenceAbsPlot  = renderPlot({
    Aseq<- seq(0.001, 0.05, length=1000)
    plot(Aseq,
       lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
#  abline(v=alpha/2, lty=2)
  #abline(v=results.EqualLR.Table()$alphas[2], lty=3)
})

```

```{r lrDifferenceAbsPlot output not used}
# not a helpful plot

### lrDifferenceAbsPlot

#plotOutput(outputId = 'lrDifferenceAbsPlot')  #, width = '40%', height='40%')

```
<!-- ## Comparing power, as alpha allocation changes -->

```{r theplotChangingAlphaBig render}
minAlphaBig = 0.002

### This plot shows how the power and cutoffs change as the alpha allocation changes
output$theplotChangingAlphaBig = renderPlot({
  cutoffsBig =   qnorm(1 - input$ChangingAlphaBig, sd = sqrt(max(vars())))
  cutoffsSmall = qnorm(1 - (alpha - input$ChangingAlphaBig), sd = sqrt(min(vars())))
  cutoffsAny = c(cutoffsSmall, cutoffsBig)
  # testthat::expect_equal(lr(cutoffsSmall, min(vars)),
  #                        lr(cutoffsBig, max(vars)) , tolerance = 1e-3)
  conditionalPowerEqualLR =   c( pow(cutoffsSmall, M.alt=1, V=vars()[1]),
                                 pow(cutoffsBig, M.alt=1, V=vars()[2]) )
  upperX = qnorm(1 - minAlphaBig, sd = sqrt(max(vars())))
  cat('upperX ', upperX, '\n')
  Cseq <- seq(lowerC, max(results.EqualLR.Table()$cutoffs), length=1000)

  plot(Cseq, ylim=c(0,1), xlim=c(0, upperX * 1.005),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
  #  Draw points for the equal alpha case.
  pow.M.alt.1 = pow(cutoffs.equalAlpha(), M.alt=1, V=vars())
  points(cutoffs.equalAlpha(), pow.M.alt.1,  pch='A', col=col.v['alpha'], cex=2)
  points(cutoffs.equalLR(), pow(cutoffs.equalLR(), M.alt=1, V=vars()), col=col.v['LR'], pch='L', cex=2)
  points(cutoffsAny, pow(cutoffsAny, M.alt=1, V=vars()), pch='?', cex=3)
  #   lines(x=c(lowerC, cutoffs()[1]), y= c(alpha,alpha) , col='black')
  # lines(x=c(lowerC, cutoffs()[2]), y= c(alpha,alpha) , col='black')
  # text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black')
  # points(cutoffs(), c(alpha, alpha), col=col.v, pch='C')
  #points(cutoffs, pow(cutoffs, M.alt=0), col=col.v, pch='C')  # null hypothesis
  # points(cutoffs(), pow.M.alt.1(), col=col.v, pch='P')
  cutoffsLR = results.EqualLR.Table()$cutoffs
  cutoffsLRSmall =   cutoffsLR[1]
  cutoffsLRBig =   cutoffsLR[2]
  for(Vn in 1:2) {
    # lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
    # lines(x=c(lowerC, cutoffsLR[Vn]), y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
    # lines(x=rep(cutoffs[Vn],2), y = c(alpha, pow.M.alt.1() [Vn]), col=col.v[Vn])
    # text(x=lowerC, y=pow.M.alt.1() [Vn], col=col.v[Vn],
    #      labels = paste('V=', vars()[Vn], '  power=',
    #                     signif(digits = 2, pow.M.alt.1() [Vn])), adj = c(0,0))

  }
  ### average
  margPowerEqualAlpha = mean(pow.M.alt.1)
  #  lines(x=c(lowerC, mean(cutoffs)), y= rep(margPowerEqualAlpha, 2), col=col.v[3])
  #lines(rep(lowerC, 2), pow.M.alt.1, col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power for equal alpha:',
                      signif(digits = 3, margPowerEqualAlpha)))

  conditionalPowerEqualLR = results.EqualLR.Table()$power
  margPowerEqualLR = mean(conditionalPowerEqualLR)
  points( cex=2, c(  cutoffsLRSmall,   cutoffsLRBig),
          conditionalPowerEqualLR, col=col.v['LR'], pch='L')
  points( cex=3, c(  cutoffsLRSmall,   cutoffsLRBig),
          conditionalPowerEqualLR, col=col.v['LR'], pch=1)
  text(x=lowerC, y=margPowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'], xpd=NA,
       labels = paste('marginal power equal LR:',
                      signif(digits = 3, margPowerEqualLR)))
  legend(x='topright', 
         legend=c(paste('Variance=', vars()[1]),
                  paste('Variance=', vars()[2]),
                  'conditional power for equal alpha',
                  'conditional power for equal LR',
                  'conditional power for other allocation'
         ),
         text.col =  c(col.v, 'black'),
         pch = c('', '', 'A', 'L', '?'),
        
  )
})


```

```{r theplotChangingAlphaBig output, echo=FALSE}
## We are FOR NOW keeping Pr(variance = vars[1] equal to 1/2).
numericInput('ChangingAlphaBig', 'Changing how much alpha is allocated to the Big variance part', value = alpha,
             min=minAlphaBig, max=alpha * 2 - minAlphaBig, step=minAlphaBig)

plotOutput('theplotChangingAlphaBig')
```


```{r numeric inputs}
div(
  numericInput('varsBig', 'variance Big', value = vars.init[2], min = vars.init[1], step=vars.step),
  numericInput('varsSmall', 'variance Small', value = vars.init[1], min = 0,
               max = vars.init[2], step=vars.step)
)
vars = reactive({c(input$varsSmall, input$varsBig)})
observeEvent(input$varsBig,
             updateNumericInput(inputId='varsSmall', max = input$varsBig))
observeEvent(input$varsSmall,
             updateNumericInput(inputId='varsBig', min = input$varsSmall))
```

```{r theplotForEquals, echo=FALSE}

#### Plot power and cutoffs for equalAlpha and equalLR cases.

output$theplotForEquals = renderPlot({
    Cseq <- seq(lowerC, max(results.EqualLR.Table()$cutoffs), length=1000)

    cutoffsLR = results.EqualLR.Table()$cutoffs
    cutoffsLRSmall =   cutoffsLR[1]
    cutoffsLRBig =   cutoffsLR[2]
    cutoffsAlpha = results.EqualAlpha.Table()$cutoffs
    cutoffsAlphaSmall =   cutoffsAlpha[1]
    cutoffsAlphaBig =   cutoffsAlpha[2]
  plot(Cseq, 
       ylim=c(0,1), xlim=c(0,   cutoffsLRBig * 1.01),
       pow(Cseq, M.alt=0, 1), 
       xlab='critical rejection cutoff, conditional on which experiment',
       ylab='power to reject alternative',
       pch='', col=col.v1)
  lines(Cseq, pow(Cseq, M.alt=1, vars()[1]), col=col.v1)
  lines(Cseq, pow(Cseq, M.alt=1, vars()[2]), col=col.v2)
  #####  LR ########
  # lines(x=c(lowerC,   cutoffsLR[1]), y= c(alpha,alpha) , col='black')
  # lines(x=c(lowerC,   cutoffsLR[2]), y= c(alpha,alpha) , col='black')
  # text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  
    conditionalPowerEqualLR = results.EqualLR.Table()$power

  for(v in 1:2) {
    lines(rep(cutoffsAlpha[v],2), 
          c(alpha, pow(cutoffsAlpha[v], M.alt=0, V=vars()[v])))
    points(cutoffsAlpha[v], 
           results.EqualAlpha.Table()$power[v], col=col.v[v], pch='A', cex=2)
    text(x= cutoffsAlpha[v] + ifelse(v==1, -0.8, -0.8) ,
         y=results.EqualAlpha.Table()$power[v],
          xpd=NA,  
         cex=1, 
         'conditional power\nequal Alpha', col=col.v[v])
    text(pos=ifelse(v==1, 3, 2), xpd=NA,  cex=1, 
         cutoffsLR[v] - 0.05*(v==2), results.EqualLR.Table()$power[v]
         + 0.05*(v==1) - 0.02*(v==2), 
          'conditional power\nequal LR', col=col.v[v])
  }

    points( cex=2, c(cutoffsLRSmall, cutoffsLRBig), 
          conditionalPowerEqualLR, col=col.v, pch='L')
  
  
  # for(Vn in 1:2) {
  #   lines(Cseq, pow(Cseq, M.alt=1, V=vars()[Vn]), lty=2, lwd=3, col=col.v[Vn])
  #   lines(x=c(lowerC, cutoffs.equalAlpha[Vn]), 
  #         y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
  #   lines(x=rep(cutoffs.equalAlpha[Vn],2), 
  #         y = c(alpha, pow.M.alt.1 [Vn]), col=col.v[Vn])
  #   text(x=lowerC, y=pow.M.alt.1 [Vn], col=col.v[Vn],
  #        labels = paste('V=', vars()[Vn], '  power=', 
  #                       signif(digits = 2, pow.M.alt.1 [Vn])), adj = c(0,0))
  #   
  # }
  ### average
  margPowerEqualAlpha = signif(digits=4, mean(results.EqualAlpha.Table()$power))
  margPowerEqualLR = signif( digits=4, mean(results.EqualLR.Table()$power))
  
  text(x=2.8, y = margPowerEqualLR + 0.07, adj = c(0,-1), xpd=NA, 'marginal power:',
       font=c(4))
  text(x=2.8, y = margPowerEqualAlpha, xpd=NA, adj = c(0,1), 
       col=col.v['alpha'], xpd=NA,
       labels = paste('equal alpha:', 
                      signif(digits = 3, margPowerEqualAlpha)))
  lines(c(0, 2.8), rep(margPowerEqualAlpha,2),  col=col.v['alpha'])
  text(x=2.8, y=margPowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v['LR'],
       labels = paste('equal LR:', 
                      signif(digits = 3, margPowerEqualLR)))
  lines(c(0, 2.8), rep(margPowerEqualLR,2),  col=col.v['LR'])

  legend(x='topright', pch = c('','','A', 'L'),
         legend=c(paste('small variance=', vars()[1]), 
                  paste('big    variance=', vars()[2]),
                  'equal alpha',
                  'equal likelihood'
         ), 
         text.col =  c(col.v1, col.v2, col.v['alpha'], col.v['LR']),
        
  )

})
```

## Comparing power, for equal alpha (A) and equal LR (L)


```{r tables and plot}
centerIt = function(S) span(style='text-align:center;  display: flex;
  align-items: center;
  justify-content: center; font-size:larger', em(S))
output$meanAlphaPower = renderText(
  paste('Marginal power = ' , signif( mean(results.EqualAlpha.Table()$power))))
output$meanLRPower = renderText(
  paste('Marginal power = ', signif(mean(results.EqualLR.Table()$power))))
output$alphaSmall.equalLR = output$alphaSmall.equalLR.dup = renderText({
  value = signif(digits=2, results.EqualLR.Table()$alpha[1])
  value
})
output$alphaBig.equalLR = renderText({
  value = signif(digits=2, results.EqualLR.Table()$alpha[2])
  value
})
fluidRow(column(5, 
                centerIt('Equal alpha'), 
                 centerIt(textOutput('meanAlphaPower')), 
                tableOutput('results.EqualAlpha'),
                centerIt('Equal LR'), 
                centerIt(textOutput('meanLRPower')),
                tableOutput('results.EqualLR')
                ),
         column(7, plotOutput('theplotForEquals')))

```

You are violating conditionality principle--
condition only on the experiment you actually did.
So for negligible gain in marginal statistical power,  you do a stupid thing.

So what happens if the small variance, more accurate instrument is picked, and the observation X is exactly at the cutoff for the same-likelihood-ratio plan with alpha = 0.05?
Well then you can claim that the P value is 0.05. But being honest, which means conditional on the experiment actually done, the data X is actually at the conditional cutoff corresponding to alpha = `r tO('alphaSmall.equalLR')`. To paraphrase,

"The coin flip could have picked a **worse** experiment. But even though it didn't, the existence of that possibility means that we can report that the strength of evidence is 0.05 not `r tO('alphaSmall.equalLR.dup')`."

And what's just as bad,  this can happen:  

"The coin flip could have picked a **better** experiment. But even though it didn't, the existence of that possibility means that we MUST report that the strength of evidence is only 0.05 not `r tO('alphaBig.equalLR')`."

CONCLUSION:  Taking seriously the mandate to choose a method that maximizes power is bad.  Always condition on the ancillary statistic when you can.  Here the ancillary statistic is which experiment was chosen.


