---
title: "Untitled"
author: "Roger Day"
date: "2026-07-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


#  neyman pearson funky demo
# violating conditionality principle



```{r echo=FALSE}

h0 = 0  # null
ha = 1 #  alternative hypothesis
vars = c(0.1, 2) ####    Small, and large. so power is large, Small.
alpha = 0.05
col.v1 = 'red'
col.v2 = 'darkgreen'
col.v = c(col.v1, col.v2, 'blue', 'orange')
names(col.v) = c('small', 'big', 'alpha', 'LR')
pow = function(C, M.alt=1, V) {
  1 - pnorm(C, M.alt, sqrt(V)) 
}
  lowerC = 0
  Cseq <- seq(lowerC,3,length=1000)

likelihood = function(h, v, x) {
   1 - pnorm( abs((x - h)/sqrt(v))  ) 
}
likelihood(h = 0, v = 1, x = qnorm(1-0.025))
lr = function(x, v) likelihood(h = ha, v , x ) / likelihood(h = h0, v, x )
```

```{r eval=FALSE}

### skip this plot
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))
abline(v=0:1)
lines(col='red',xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[2])))
plot(xseq<- seq(lowerC, 3, length=1000),  log(lr(xseq, vars[1])))

```

If we know the variance is $V$, and the observed value is $X$,
the likelihood function as a function of the mean $M$ is 
$$
L(M)~=~dnorm(X,M,\sqrt{V})
$$
For rejection critical value $C$ > 0, 
the power to reject $M = 0$ if the true mean is $M_{alt}$ is

$$
Power(C)~=~  (1 - pnorm(C,M_{alt},\sqrt{V}) )
$$
The likelihood ratios at the two critical values are
```{r echo=FALSE}
  C95  = qnorm(1 - alpha, sd = sqrt(vars))

lr(C95, vars)

```

If we raise the cutoff for the Bigger variance, and lower it for the Smaller variance, watch what happens.

```{r}
Aseq<- seq(0.001, 0.05, length=1000)
lrBigValues = numeric(0)
lrSmallValues = numeric(0)
C95BigValues = numeric(0)
C95SmallValues = numeric(0)
lrDifference = function(alphaBig){
  alphaSmall = 2*alpha  - alphaBig # because they average to alpha.
  C95Big = qnorm(1 - alphaBig, sd = sqrt(max(vars)))
  C95Small = qnorm(1 - alphaSmall, sd = sqrt(min(vars)))
  C95BigValues =  append(C95BigValues, C95Big)
  C95SmallValues =  append(C95SmallValues, C95Small)
  #print(c( lr(C95Big, v = max(vars)) , lr(C95Small, v = min(vars)) ) )
  lrBig = lr(C95Big, v = max(vars)) 
  lrSmall = lr(C95Small, v = min(vars))
  lrBigValues =  append(lrBigValues, lrBig)
  lrSmallValues =  append(lrSmallValues, lrSmall)
  lrBig - lrSmall
}
lrDifferenceAbs = function(alphaBig) abs(lrDifference(alphaBig))
# plot(Aseq,
#      lrDifferenceAbs(Aseq) , xlab='alpha Big', ylab='lrDifferenceAbs')
optimalAlphaBig = optimize(lrDifferenceAbs, 
                           maximum = FALSE, interval = range(Aseq)) $ minimum
whichOptimal = print( which(abs(optimalAlphaBig - Aseq) < 1e-4) [1] )
optimalAlphaSmall = 2*alpha - optimalAlphaBig
C95Big =   qnorm(1 - optimalAlphaBig, sd = sqrt(max(vars)))
C95Small = qnorm(1 - optimalAlphaSmall, sd = sqrt(min(vars)))
testthat::expect_equal(lr(C95Small, min(vars)), 
                       lr(C95Big, max(vars)) , tolerance = 1e-3)
conditionalPowerEqualLR =   c( pow(C95Small, M.alt=1, V=vars[1]), 
                               pow(C95Big, M.alt=1, V=vars[2]) )

averagePowerEqualLR = mean(conditionalPowerEqualLR)
  

```


```{r echo=FALSE}
### conditional power

output$theplot = renderPlot({
  plot(Cseq, ylim=c(0,1), xlim=c(0, C95Big * 1.01),
       pow(Cseq, M.alt=0, 1), xlab='critical rejection cutoff', ylab='power',
       pch='', col=col.v1)
  lines(x=c(lowerC, C95[1]), y= c(alpha,alpha) , col='black')
  lines(x=c(lowerC, C95[2]), y= c(alpha,alpha) , col='black')
  text(x=lowerC, y=alpha, labels = paste('alpha=', alpha), adj = c(0,0), col='black') 
  points(C95, c(alpha, alpha), col=col.v, pch='C')
  #points(C95, pow(C95, M.alt=0), col=col.v, pch='C')
  pow.M.alt.1 = pow(C95, M.alt=1, V=vars)
  points(C95, pow.M.alt.1, col=col.v, pch='P')
  
  
  for(Vn in 1:2) {
    lines(Cseq, pow(Cseq, M.alt=1, V=vars[Vn]), lty=2, lwd=3, col=col.v[Vn])
    lines(x=c(lowerC, C95[Vn]), y= rep(pow.M.alt.1 [Vn], 2), col=col.v[Vn])
    lines(x=rep(C95[Vn],2), y = c(alpha, pow.M.alt.1 [Vn]), col=col.v[Vn])
    text(x=lowerC, y=pow.M.alt.1 [Vn], col=col.v[Vn],
         labels = paste('V=', vars[Vn], '  power=', 
                        signif(digits = 2, pow.M.alt.1 [Vn])), adj = c(0,0))
    
  }
  ### average
  averagePowerEqualAlpha = mean(pow.M.alt.1)
#  lines(x=c(lowerC, mean(C95)), y= rep(averagePowerEqualAlpha, 2), col=col.v[3])
  lines(rep(lowerC, 2), pow.M.alt.1, col=col.v[3])
  text(x=lowerC, y=mean(pow.M.alt.1), xpd=NA, adj = c(0,1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal alpha:', 
                      signif(digits = 6, averagePowerEqualAlpha)))

  
  points( cex=2, c(C95Small, C95Big), 
          conditionaPowerEqualLR, col=col.v, pch='L')
  text(x=lowerC, y=averagePowerEqualLR, xpd=NA, adj = c(0,-1), col=col.v[3], xpd=NA,
       labels = paste('marginal power equal LR:', 
                      signif(digits = 6, averagePowerEqualLR)))
  
  
  legend(x='topright', 
         legend=c(paste('Variance=', vars[1]), 
                  paste('Variance=', vars[2]), 
                  'equal alpha',
                  'equal LR'
         ), 
         text.col =  col.v
  )
})

output$likrat = renderPlot({
  plot(c(C95SmallValues, C95BigValues), 
       lr)
})
plotOutput('theplot')
```



```{r}
cutoffs = qnorm(0.95, 0, sqrt(vars))
```

If V=1, the cutoff for alpha=0.05 is `r cutoffs[1] `.
But if V=2, the cutoff for alpha=0.05 is `r cutoffs[2] `.
The likelihood ratios are, respectively, 

`r lr(cutoffs[1], vars[1])` and `r lr(cutoffs[2], vars[2])`

The average power, conditioning on which variance is chosen:
The equalizer for the likelihood ratio is when  
AlphaBig = optimalAlphaBig = `r optimalAlphaBig` and  
AlphaSmall = optimalAlphaSmall = `r optimalAlphaSmall`.  
The critical value for the small variance is now `r C95Small`,
and for the large variance it's `r C95Big`.






